---
title: "R Workflow Scholier et al.,"
format: html
embed-resources: true
editor: visual
toc: true
code-fold: true
code-summary: "Show the code"
code-tools: true
---

## Introduction

This Quarto file documents the R worflow used to analyse the data in the paper 'Effects of past and present habitat on the gut microbiota of a wild rodent' written by Tiffany Scholier, Anton Lavrinienko, Eva R. Kallio, Phillip C. Watts & Tapio Mappes.

## Legend

Throughout the code, the past habitat/site of origin is referred to as *sitetr1* and the present habitat/site of transfer is referred to as *sitetr2*. Urban and rural sites are abbreviated by *urb* and *cont* (control), respectively. For example, the term *urbcont* refers to animals that were translocated from urban forest sites to rural forest sites. Overall, time point 1 refers to pre-transfer samples and time point 3 refers to post-transfer samples. 

## Loading necessary R packages

```{r}
#| output: false
library(phyloseq) #working with phyloseq objects
library(dplyr)    #for %>% function
library(qiime2R)  #for read_qza function
library(ggplot2)  #creating plots
library(ggpubr)   #for ggpaired function
library(vegan)    #for adonis2 function
library(lmerTest) #for lmer function
library(rstatix)  #for add_significance function

packageVersion("phyloseq") #1.44.0
packageVersion("dplyr")    #1.1.3
packageVersion("qiime2R")  #0.99.6
packageVersion("ggplot2")  #3.4.3
packageVersion("ggpubr")   #0.6.0
packageVersion("vegan")    #2.6.4
packageVersion("lmerTest") #3.1.3
packageVersion("rstatix")  #0.7.2
```

## Go to the directory

```{r}
setwd("~/Documents/R/RT")
```

## Preparing all necessary R files

```{r}
#| output: false

#DATAFRAME FILES

meta<-read.table("metadata_quarto.txt", header=T)
meta$treatment <- factor(meta$treatment, levels = c("contcont", "urbcont", "conturb", "urburb"))
meta.mom<-meta %>% filter(type == "mom")
alpharich<-read_qza("observed_features_vector.qza")$data %>% tibble::rownames_to_column("sampleid")
shannon<-read_qza("shannon_vector.qza")$data %>% tibble::rownames_to_column("sampleid")
faith<-read_qza("faith_pd_vector.qza")$data %>% tibble::rownames_to_column("sampleid")

diff.13<-read.table("differences_1_to_3_from_qiime.txt", header=T)
diff.13$treatment <- factor(diff.13$treatment, levels = c("contcont", "urbcont", "conturb", "urburb"))
diff.13$sitetr1<-as.factor(diff.13$sitetr1)
diff.13$sitetr2<-as.factor(diff.13$sitetr2)
diff.13$change<-as.factor(diff.13$change)

dist.13<-read.table("distances_1_to_3_from_qiime.txt", header=T)
dist.13$treatment <- factor(dist.13$treatment, levels = c("contcont", "conturb", "urbcont", "urburb"))
dist.13$sitetr1<-as.factor(dist.13$sitetr1)
dist.13$sitetr2<-as.factor(dist.13$sitetr2)
dist.13$change<-as.factor(dist.13$change)
dist.13.natives<-subset(dist.13, treatment=="urburb" | treatment == "contcont")

metadata<-meta%>% left_join(alpharich)%>% left_join(shannon)%>% left_join(faith)
metadata$sitetr1<-as.factor(metadata$sitetr1)
metadata$sitetr2<-as.factor(metadata$sitetr2)
metadata$time<-as.factor(metadata$time)
metadata$treatment <- factor(metadata$treatment, levels = c("contcont", "conturb", "urbcont", "urburb"))

metadata.mom<-metadata %>% filter(type == "mom")
metadata.mom.time1 <- metadata.mom %>% filter(time == "1")
metadata.mom.time3 <- metadata.mom %>% filter(time == "3")
metadata.mom.fromurban <- metadata.mom %>% filter(sitetr1 == "urban")
metadata.mom.fromcontrol <- metadata.mom %>% filter(sitetr1 == "control")

metadata.pup<-metadata %>% filter(type == "pup")
metadata.pup$geneticcluster<-as.factor(metadata.pup$geneticcluster)

########################################################################
#RARIFIED PHYLOSEQ

psrt<-qza_to_phyloseq(
  features="rt_table16s_filtermin10freq_rar30671.qza",taxonomy="rt_taxonomysilva.qza",metadata = "metadata_quarto.txt",tree = "rt_rooted-tree.qza")
levels(sample_data(psrt)$sitetr1) <- c("control", "urban")
levels(sample_data(psrt)$sitetr2) <- c("control", "urban")
levels(sample_data(psrt)$treatment) <- c("contcont", "conturban", "urbancont","urbanurban")

psrt.mom = psrt %>% subset_samples(type == "mom")
metadata.psrt.mom <- as(sample_data(psrt.mom), "data.frame")
psrt.mom.time3 = subset_samples(psrt.mom, time != "1")
psrt.mom.time1 = subset_samples(psrt.mom, time == "1")

psrt.pup = psrt %>% subset_samples(type == "pup")
metadata.psrt.pup <- as(sample_data(psrt.pup), "data.frame")

psrt.mom3.pup = subset_samples(psrt, time != "1")
metadata.psrt.mom3.pup <- as(sample_data(psrt.mom3.pup), "data.frame")

psrt.matched = psrt.mom3.pup %>% subset_samples(geneticcluster_2 %in% c("C1", "C5", "C8","C9","C10","C12","C14","C15","C17","C18","C23","C24","C26"))
metadata.matched <- as(sample_data(psrt.matched), "data.frame")

```

## Recreate Figure 1

**Experimental design of the reciprocal translocation (RT) experiment on wild bank voles**

This figure was made with PowerPoint and GIS software.

## Recreate Figure 2

**Change in the alpha diversity of the gut microbiota of bank voles in terms of past habitat**

```{r}
#| fig-width: 7
#| fig-height: 5

plot.asv.13.urban <- ggpaired(metadata.mom.fromurban, 
                                  x = "time", y = "observed_features", 
                                  order = c("1", "3"),color = "black",fill= "#f00078",
                                  line.color = "gray", line.size = 0.4,id = "ind",
                                  ylab = "ASV richness", xlab = "")+
  ggtitle("Urban")+
  scale_x_discrete(labels=c("1" = "Pre-transfer", "3" = "Post-transfer"))+
  scale_y_continuous(limits = c(150,550))+theme_minimal()

plot.asv.13.rural <- ggpaired(metadata.mom.fromcontrol, 
                              x = "time", y = "observed_features", 
                              order = c("1", "3"),color = "black",fill= "#693aa3",
                              line.color = "gray", line.size = 0.4,id = "ind",
                              ylab = "ASV richness", xlab = "")+
  ggtitle("Rural")+
  scale_x_discrete(labels=c("1" = "Pre-transfer", "3" = "Post-transfer"))+
  scale_y_continuous(limits = c(100,350))+theme_minimal()

plot.asv.t3<-ggplot(data=metadata.mom.time3, aes(x=sitetr1, y=observed_features, fill=sitetr1))+
  geom_boxplot() +
  theme(legend.position="right")+
  labs(title= "Post-transfer gut microbiota", x= "",y = "")+
  scale_x_discrete(labels=c("Rural","Urban"))+
  ylab("ASV richness")+
  scale_fill_manual(values=c("#693aa3","#f00078"),name=c(""),
                    labels=c("Rural","Urban"))+
  geom_signif(y_position = c(550), xmin = c(1), xmax = c(2),annotation = c("*"), tip_length = 0)+
  theme_minimal()

plot.change.asv<-ggarrange(plot.asv.13.rural, plot.asv.13.urban,
                           common.legend = FALSE,
                           legend = "right",
                           labels = c("B", "C"),
                           ncol = 1, nrow = 2, align="v"+ theme_void())

plot.asv<-ggarrange(plot.asv.t3, plot.change.asv, 
                           common.legend = FALSE,
                           legend = "none",
                           labels = c("A", ""),
                           ncol = 2, nrow = 1, align="v"+ theme_void())
plot.asv #(6 x 6)
```

## Recreate Figure 3

**Post-transfer gut microbiota compositions of bank voles in terms of past and present habitat**

Panels were either made with R (A and B) or PowerPoint sotftware (C and D).

```{r}
#| fig-width: 10
#| fig-height: 4

set.seed(123)
bray_dist.mom.time3 = phyloseq::distance(psrt.mom.time3, method="bray")
set.seed(123)
cap.mom.3.bray <- ordinate(physeq = psrt.mom.time3, method = "CAP",distance = bray_dist.mom.time3,formula = ~ sitetr1+sitetr2)
plot.mom.3.bray<-plot_ordination(psrt.mom.time3, cap.mom.3.bray,type = "samples",color="treatment")+geom_point(size = 3)+
  scale_color_manual(values=c("#693aa3","#7777ff","#ffc1e0","#f00078"), name=(""),
                     labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  ggtitle("Adults")+xlab("CAP Axis.1  [8%]")+ylab("CAP Axis.2  [3.8%]")+theme_minimal()

set.seed(123)
bray_dist.pup.all = phyloseq::distance(psrt.pup, method="bray")
set.seed(123)
cap.pup.all.bray <- ordinate(physeq = psrt.pup, method = "CAP",distance = bray_dist.pup.all,formula = ~ sitetr1+sitetr2)
plot.pup.cap.bray<-plot_ordination(psrt.pup, cap.pup.all.bray,type = "samples",color="treatment")+geom_point(size = 3)+
  scale_color_manual(values=c("#693aa3","#7777ff","#ffc1e0","#f00078"), name=(""),
                     labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  ggtitle("Offspring")+xlab("CAP Axis.1  [3.9%]")+ylab("CAP Axis.2  [1.9%]")+theme_minimal()

plot.cap.fig3<-ggarrange(plot.mom.3.bray, plot.pup.cap.bray,
                           common.legend = TRUE,
                           legend = "right",
                           labels = c("A", "B"),
                           ncol = 2, nrow = 1, align="v"+ theme_void())
plot.cap.fig3 # 10 x 12
```

## Recreate Table S1

**Differences in beta diversity between the pre-transfer gut microbiota of urban and rural bank voles**

```{r}
#| output: false
#Preparation (calculating distances + generating new metadata file)
set.seed(123)
bray_dist.mom.time1 = phyloseq::distance(psrt.mom.time1, method="bray")
set.seed(123)
jacc_dist.mom.time1 = phyloseq::distance(psrt.mom.time1, method="jaccard")
set.seed(123)
wunifrac_dist.mom.time1 = phyloseq::distance(psrt.mom.time1, method="wunifrac")
set.seed(123)
uunifrac_dist.mom.time1 = phyloseq::distance(psrt.mom.time1, method="uunifrac")
metadata.psrt.time1 <- as(sample_data(psrt.mom.time1), "data.frame")
```
### Bray-Curtis metric
```{r}
set.seed(123)
adonis.mom.1<-adonis2(bray_dist.mom.time1~sitetr1, data=metadata.psrt.time1, by="margin",perm=9999)
adonis.mom.1 #R2 = 9%, p-value = 3e-04
set.seed(123)
mod.mom.1 <- betadisper(bray_dist.mom.time1, metadata.psrt.time1$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.1) #No heterogeneity
boxplot(mod.mom.1)
```
### Jaccard metric
```{r}
set.seed(123)
adonis.mom.1.jacc<-adonis2(jacc_dist.mom.time1~sitetr1, data=metadata.psrt.time1, by="margin", perm=9999)
adonis.mom.1.jacc #R2 = 6.8%, p-value = 3e-04
set.seed(123)
mod.mom.1.jacc <- betadisper(jacc_dist.mom.time1, metadata.psrt.time1$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.1.jacc) #No heterogeneity
boxplot(mod.mom.1.jacc)
```
### Weighted Unifrac metric
```{r}
set.seed(123)
adonis.mom.1.wu<-adonis2(wunifrac_dist.mom.time1~sitetr1, data=metadata.psrt.time1, by="margin", perm=9999)
adonis.mom.1.wu #p-value > 0.05
set.seed(123)
mod.mom.1.wu <- betadisper(wunifrac_dist.mom.time1, metadata.psrt.time1$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.1.wu) #No heterogeneity
boxplot(mod.mom.1.wu)
```
### Unweighted Unifrac metric
```{r}
set.seed(123)
adonis.mom.1.uu<-adonis2(uunifrac_dist.mom.time1~sitetr1, data=metadata.psrt.time1, by="margin", perm=9999)
adonis.mom.1.uu #R2 = 7.7%, p-value = 2e-04
set.seed(123)
mod.mom.1.uu <- betadisper(uunifrac_dist.mom.time1, metadata.psrt.time1$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.1.uu) #Heterogeneity!
boxplot(mod.mom.1.uu) #Higher variance in rural samples in comparison to urban samples
```



## Recreate Table S2

**Differences in alpha diversity between the pre-transfer gut microbiota of urban and rural bank voles**

```{r}
#| output: false
#Test normality
shapiro.test(metadata.mom.time1$observed_features) #p-value = 0.264 = NORMAL DISTRIBUTION
shapiro.test(metadata.mom.time1$shannon_entropy) #p-value = 0.458 = NORMAL DISTRIBUTION
shapiro.test(metadata.mom.time1$faith_pd) #p-value = 0.6019 = NORMAL DISTRIBUTION
#Normality = OK -> use linear models
```
### ASV richness
```{r}
set.seed(123)
model_asv_t1<-lm(observed_features~sitetr1, data=metadata.mom.time1)
summary(model_asv_t1)

```
### Shannon diversity
```{r}
set.seed(123)
model_sh_t1<-lm(shannon_entropy~sitetr1, data=metadata.mom.time1)
summary(model_sh_t1)
```
### Faith's phylogenetic diversity
```{r}

set.seed(123)
model_faith_t1<-lm(faith_pd~sitetr1, data=metadata.mom.time1)
summary(model_faith_t1)
```


## Recreate Table S3

**Temporal change in the alpha diversity of the gut microbiota of bank voles within each experimental group**

```{r}
#| output: false
#Order the dataframes (same individuals should we listed together)
metadata.mom.1.3 <- metadata.mom %>% arrange(ind)
```
```{r}
#| output: false
# Test normality of the differences
shapiro.test(diff.13$asv_1_to_3) #p-value = 0.8119 = NORMAL DISTRIBUTION
shapiro.test(diff.13$shannon_1_to_3) #p-value = 0.9922 = NORMAL DISTRIBUTION
shapiro.test(diff.13$faith_1_to_3) #p-value = 0.5452 = NORMAL DISTRIBUTION
```
### ASV richness
```{r}
#Normality = OK -> Use t-tests
stat.test.13<-metadata.mom.1.3 %>%
  group_by(treatment) %>%
  t_test(observed_features ~ time, paired=TRUE,alternative = "two.sided") %>%
  adjust_pvalue(method = "fdr") %>% add_significance("p.adj") 
stat.test.13

bp <- ggpaired(metadata.mom.1.3, x = "time", y = "observed_features", order = c("1", "3"), color="treatment",
               line.color = "gray", line.size = 0.4,id = "ind",ylab = "ASV richness", xlab = "")+ 
    scale_x_discrete(labels=c("1" = "Pre-transfer", "3" = "Post-transfer"))+
    facet_wrap(~ treatment, nrow=2,ncol=2)
bp
```
### Shannon diversity
```{r}
stat.test.13.shannon<-metadata.mom.1.3 %>%
  group_by(treatment) %>%
  t_test(shannon_entropy ~ time, paired=TRUE,alternative = "two.sided") %>%
  adjust_pvalue(method = "fdr") %>%add_significance("p.adj") 
stat.test.13.shannon

bp.sh <- ggpaired(metadata.mom.1.3, x = "time", y = "shannon_entropy", order = c("1", "3"), color="treatment",
                  line.color = "gray", line.size = 0.4,id = "ind",ylab = "Shannon diversity", xlab = "")+ 
    scale_x_discrete(labels=c("1" = "Pre-transfer", "3" = "Post-transfer"))+ 
    facet_wrap(~ treatment, nrow=2,ncol=2)
bp.sh
```
### Faith's phylogenetic diversity
```{r}
stat.test.13.faith<-metadata.mom.1.3 %>%
  group_by(treatment) %>%
  t_test(faith_pd ~ time, paired=TRUE,alternative = "two.sided") %>%
  adjust_pvalue(method = "fdr") %>%add_significance("p.adj") 
stat.test.13.faith

bp.pd <- ggpaired(metadata.mom.1.3, x = "time", y = "faith_pd", order = c("1", "3"), color="treatment",
                  line.color = "gray", line.size = 0.4,id = "ind",ylab = "Faith's PD", xlab = "")+ 
    scale_x_discrete(labels=c("1" = "Pre-transfer", "3" = "Post-transfer"))+ 
    facet_wrap(~ treatment, nrow=2,ncol=2)
bp.pd
```




## Recreate Table S4

**Pairwise differences in alpha diversity (temporal alpha diversity change) between the gut microbiota of bank voles belonging to different experimental groups**

### ASV richness
```{r}
#Normality = OK (see 'Recreate Table S2')-> Use linear models
model.13.asv <- lm(asv_1_to_3 ~ sitetr1*sitetr2, data = diff.13)
summary(model.13.asv) 
model.13.asv.2 <- lm(asv_1_to_3 ~ sitetr1+sitetr2, data = diff.13)
summary(model.13.asv.2) 
model.13.asv.3 <- lm(asv_1_to_3 ~ sitetr1, data = diff.13)
summary(model.13.asv.3) 
```
### Shannon diversity
```{r}
model.13.shannon <- lm(shannon_1_to_3 ~ sitetr1*sitetr2, data = diff.13)
summary(model.13.shannon)
model.13.shannon.2 <- lm(shannon_1_to_3 ~ sitetr1+sitetr2, data = diff.13)
summary(model.13.shannon.2)
```
### Faith's phylogenetic diversity
```{r}
model.13.faith <- lm(faith_1_to_3 ~ sitetr1*sitetr2, data = diff.13)
summary(model.13.faith)
model.13.faith.2 <- lm(faith_1_to_3 ~ sitetr1+sitetr2, data = diff.13)
summary(model.13.faith.2)
```
### All metrics in terms of change
```{r}
model.13.asv.change <- lm(asv_1_to_3 ~ change, data = diff.13)
summary(model.13.asv.change)
model.13.shannon.change <- lm(shannon_1_to_3 ~ change, data = diff.13)
summary(model.13.shannon.change)
model.13.faith.change <- lm(faith_1_to_3 ~ change, data = diff.13)
summary(model.13.faith.change)
```



## Recreate Table S5

**Differences in alpha diversity between the post-transfer gut microbiota of bank voles belonging to different experimental groups**

```{r}
#| output: false
#Test Normality
shapiro.test(metadata.mom.time3$observed_features) #p-value = 0.0513 = NORMAL DISTRIBUTION
shapiro.test(metadata.mom.time3$shannon_entropy)   #p-value = 0.7244 = NORMAL DISTRIBUTION
shapiro.test(metadata.mom.time3$faith_pd)          #p-value = 0.9103 = NORMAL DISTRIBUTION
#Normality = OK -> use linear models
```
### ASV richness
```{r}
model_asv_t3<-lm(observed_features~sitetr1*sitetr2, data=metadata.mom.time3)
summary(model_asv_t3)
model_asv_t3<-lm(observed_features~sitetr1+sitetr2, data=metadata.mom.time3)
summary(model_asv_t3)
model_asv_t3<-lm(observed_features~sitetr1, data=metadata.mom.time3)
summary(model_asv_t3) #p-value = 0.02
```
### Shannon diversity
```{r}
model_shannon_t3<-lm(shannon_entropy~sitetr1*sitetr2, data=metadata.mom.time3)
summary(model_shannon_t3)
model_shannon_t3<-lm(shannon_entropy~sitetr1+sitetr2, data=metadata.mom.time3)
summary(model_shannon_t3)
model_shannon_t3<-lm(shannon_entropy~sitetr1, data=metadata.mom.time3)
summary(model_shannon_t3) #p-value = 0.0538
```
### Faith's phylogenetic diversity
```{r}
model_faith_t3<-lm(faith_pd~sitetr1*sitetr2, data=metadata.mom.time3)
summary(model_faith_t3)
model_faith_t3<-lm(faith_pd~sitetr1+sitetr2, data=metadata.mom.time3)
summary(model_faith_t3)
model_faith_t3<-lm(faith_pd~sitetr1, data=metadata.mom.time3)
summary(model_faith_t3) #p-value = 0.105
```


## Recreate Table S6

**Pairwise distances in beta diversity (temporal beta diversity change) between the gut microbiota of bank voles belonging to different experimental groups**

```{r}
#| output: false
#Test normality
shapiro.test(dist.13$bray_1_to_3) #p-value = 0.4519 = NORMAL DISTRIBUTION
shapiro.test(dist.13$jacc_1_to_3) #p-value = 0.4296 = NORMAL DISTRIBUTION
shapiro.test(dist.13$wunifrac_1_to_3) #p-value = 0.01187 = NOT NORMAL
#This needs to be adjusted with a log transformation
dist.13$log_wunifrac_1_to_3 <- log(dist.13$wunifrac_1_to_3)
shapiro.test(dist.13$log_wunifrac_1_to_3) #p-value = 0.3476 = NORMAL DISTRIBUTION
shapiro.test(dist.13$unwunifrac_1_to_3) #0.07002 = NORMAL DISTRIBUTION
```
### Bray-Curtis metric
```{r}
lm.aov.13.bray <- lm(bray_1_to_3 ~ sitetr1*sitetr2, data = dist.13)
summary(lm.aov.13.bray)
lm.aov.13.bray.2 <- lm(bray_1_to_3 ~ sitetr1+sitetr2, data = dist.13)
summary(lm.aov.13.bray.2) 
lm.aov.13.bray.3 <- lm(bray_1_to_3 ~ sitetr1, data = dist.13)
summary(lm.aov.13.bray.3) #p-value = 0.107
```
### Jaccard metric
```{r}
lm.aov.13.jacc <- lm(jacc_1_to_3 ~ sitetr1*sitetr2, data = dist.13)
summary(lm.aov.13.jacc) 
lm.aov.13.jacc.2 <- lm(jacc_1_to_3 ~ sitetr1+sitetr2, data = dist.13)
summary(lm.aov.13.jacc.2) 
```
### Weighted Unifrac metric
```{r}
lm.aov.13.wunifrac <- lm(log_wunifrac_1_to_3 ~ sitetr1*sitetr2, data = dist.13)
summary(lm.aov.13.wunifrac) 
lm.aov.13.wunifrac.2 <- lm(log_wunifrac_1_to_3 ~ sitetr1+sitetr2, data = dist.13)
summary(lm.aov.13.wunifrac.2) 
lm.aov.13.wunifrac.3 <- lm(log_wunifrac_1_to_3 ~ sitetr1, data = dist.13)
summary(lm.aov.13.wunifrac.3) #p-value (origin) = 0.0125 
```
### Unweighted Unifrac metric
```{r}
lm.aov.13.unwunifrac <- lm(unwunifrac_1_to_3 ~ sitetr1*sitetr2, data = dist.13)
summary(lm.aov.13.unwunifrac)
lm.aov.13.unwunifrac.2 <- lm(unwunifrac_1_to_3 ~ sitetr1+sitetr2, data = dist.13)
summary(lm.aov.13.unwunifrac.2) #p-value > 0.05
```


### All metrics in terms of change
```{r}
#All metrics in terms of change
lm.aov.13.bray.change <- lm(bray_1_to_3 ~ change, data = dist.13)
summary(lm.aov.13.bray.change) #p-value = 0.0757
lm.aov.13.jacc.change <- lm(jacc_1_to_3 ~ change, data = dist.13)
summary(lm.aov.13.jacc.change) #p-value = 0.245
lm.aov.13.wu.change <- lm(log_wunifrac_1_to_3 ~ change, data = dist.13)
summary(lm.aov.13.wu.change)   #p-value = 0.311
lm.aov.13.uu.change <- lm(unwunifrac_1_to_3 ~ change, data = dist.13)
summary(lm.aov.13.uu.change)   #p-value = 0.991

plot(bray_1_to_3~treatment, data=dist.13)
plot(jacc_1_to_3~treatment, data=dist.13)
plot(log_wunifrac_1_to_3~treatment, data=dist.13)
plot(unwunifrac_1_to_3~treatment, data=dist.13)
```


## Recreate Table S7

**Differences in beta diversity between the post-transfer gut microbiota of bank voles belonging to different experimental groups**

```{r}
#| output: false
#Preparation (calculating distances + generating new metadata file)
set.seed(123)
bray_dist.mom.time3 = phyloseq::distance(psrt.mom.time3, method="bray")
set.seed(123)
jacc_dist.mom.time3 = phyloseq::distance(psrt.mom.time3, method="jaccard")
set.seed(123)
wunifrac_dist.mom.time3 = phyloseq::distance(psrt.mom.time3, method="wunifrac")
set.seed(123)
uunifrac_dist.mom.time3 = phyloseq::distance(psrt.mom.time3, method="uunifrac")
metadata.psrt.time3 <- as(sample_data(psrt.mom.time3), "data.frame")
```
### Bray-Curtis metric
```{r}
set.seed(123)
adonis.mom.3<-adonis2(bray_dist.mom.time3~sitetr1*sitetr2, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3          #p-value > 0.05

set.seed(123)
adonis.mom.3.2<-adonis2(bray_dist.mom.time3~sitetr1+sitetr2, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.2 

set.seed(123)
adonis.mom.3.3<-adonis2(bray_dist.mom.time3~sitetr1, data=metadata.psrt.time3, by="margin",perm=9999)
adonis.mom.3.3 #p = 0.078
set.seed(123)
mod.mom.bray.1 <- betadisper(bray_dist.mom.time3, metadata.psrt.time3$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.bray.1) #Heterogeneity!
boxplot(mod.mom.bray.1) #Urban > Rural

set.seed(123)
adonis.mom.3.4<-adonis2(bray_dist.mom.time3~sitetr2, data=metadata.psrt.time3, by="margin",perm=9999)
adonis.mom.3.4 #SIGNIFICANT p = 0.0381 
set.seed(123)
mod.mom.bray.2 <- betadisper(bray_dist.mom.time3, metadata.psrt.time3$sitetr2, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.bray.2) #No heterogeneity
```
### Jaccard metric
```{r}
set.seed(123)
adonis.mom.3.jacc<-adonis2(jacc_dist.mom.time3~sitetr1*sitetr2, data=metadata.psrt.time3, by="margin",perm=9999)
adonis.mom.3.jacc     #p-value > 0.05

set.seed(123)
adonis.mom.3.2.jacc<-adonis2(jacc_dist.mom.time3~sitetr1+sitetr2, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.2.jacc 

set.seed(123)
adonis.mom.3.3.jacc<-adonis2(jacc_dist.mom.time3~sitetr1, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.3.jacc #p = 0.0948
set.seed(123)
mod.mom.jacc.1 <- betadisper(jacc_dist.mom.time3, metadata.psrt.time3$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.jacc.1) #Heterogeneity!
boxplot(mod.mom.jacc.1) #Urban > Rural

set.seed(123)
adonis.mom.3.4.jacc<-adonis2(jacc_dist.mom.time3~sitetr2, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.4.jacc #SIGNIFICANT p = 0.0334
set.seed(123)
mod.mom.jacc.2 <- betadisper(jacc_dist.mom.time3, metadata.psrt.time3$sitetr2, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.jacc.2) #No heterogeneity
```
### Weighted Unifrac metric
```{r}
set.seed(123)
adonis.mom.3.wunifrac<-adonis2(wunifrac_dist.mom.time3~sitetr1*sitetr2, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.wunifrac #p-value > 0.05

set.seed(123)
adonis.mom.3.2.wunifrac<-adonis2(wunifrac_dist.mom.time3~sitetr1+sitetr2, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.2.wunifrac 

set.seed(123)
adonis.mom.3.3.wunifrac<-adonis2(wunifrac_dist.mom.time3~sitetr1, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.3.wunifrac #p = 0.122
set.seed(123)
mod.mom.wunifrac.1 <- betadisper(wunifrac_dist.mom.time3, metadata.psrt.time3$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.wunifrac.1) #Heterogeneity!
boxplot(mod.mom.wunifrac.1) #Urban > Rural

set.seed(123)
adonis.mom.3.4.wunifrac<-adonis2(wunifrac_dist.mom.time3~sitetr2, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.4.wunifrac #p =  0.1722
set.seed(123)
mod.mom.wunifrac.2 <- betadisper(wunifrac_dist.mom.time3, metadata.psrt.time3$sitetr2, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.wunifrac.2) #No heterogeneity
```
### Unweighted Unifrac metric
```{r}
set.seed(123)
adonis.mom.3.uunifrac<-adonis2(uunifrac_dist.mom.time3~sitetr1*sitetr2, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.uunifrac #p-value > 0.05

set.seed(123)
adonis.mom.3.2.uunifrac<-adonis2(uunifrac_dist.mom.time3~sitetr1+sitetr2, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.2.uunifrac 

set.seed(123)
adonis.mom.3.3.uunifrac<-adonis2(uunifrac_dist.mom.time3~sitetr1, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.3.uunifrac #p = 0.122
set.seed(123)
mod.mom.uunifrac.1 <- betadisper(uunifrac_dist.mom.time3, metadata.psrt.time3$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.uunifrac.1) #No heterogeneity

set.seed(123)
adonis.mom.3.4.uunifrac<-adonis2(uunifrac_dist.mom.time3~sitetr2, data=metadata.psrt.time3, by="margin", perm=9999)
adonis.mom.3.4.uunifrac #p =  0.1722
set.seed(123)
mod.mom.uunifrac.2 <- betadisper(uunifrac_dist.mom.time3, metadata.psrt.time3$sitetr2, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
anova(mod.mom.uunifrac.2) #No heterogeneity
```


## Recreate Table S8

**Differences in alpha diversity and beta diversity between the post-transfer gut microbiota of offspring belonging to different experimental groups**

The offspring data consists of dependent data points since multiple siblings are included into the same dataset. For analyses concerning alpha diversity, we can simply include genetic cluster (i.e., siblings belong to the same group) as a random effect in lmer models. For beta diversity analyses, there are no straightforward methods to include random effects within adonis2 models. To deal with the dependence in our data, we coded a loop (n iterations = 1000) where 1) one sibling per genetic cluster gets randomly selected (33 individuals are left), 2) the adonis2 output gets computed based upon this reduced dataset and 3) the average value for all output generated over 1000 runs gets calculated.

<span style="color:red;">The same output is generated by the loop when 1000 or 10 iterations are used, so the workflow in this quarto document contains the lowest number of iterations for convenience and efficiency.</span>

```{r}
#| output: false
shapiro.test(metadata.pup$observed_features)     #p-value = 0.0003368
metadata.pup$observed_features_log <-log(metadata.pup$observed_features)
shapiro.test(metadata.pup$observed_features_log) #p-value = 0.2536
shapiro.test(metadata.pup$shannon_entropy)       #p-value = 0.1673
shapiro.test(metadata.pup$faith_pd)              #p-value = 0.09069
```
### ASV richness
```{r}
set.seed(123)
model3<-lmer(observed_features_log~sitetr1*sitetr2 + (1|geneticcluster), data=metadata.pup)
summary(model3)

set.seed(123)
model3.1<-lmer(observed_features_log~sitetr1+sitetr2 + (1|geneticcluster), data=metadata.pup)
summary(model3.1)
```
### Shannon diversity
```{r}
set.seed(123)
model3.sh<-lmer(shannon_entropy~sitetr1*sitetr2 + (1|geneticcluster), data=metadata.pup)
summary(model3.sh)

set.seed(123)
model3.sh.1<-lmer(shannon_entropy~sitetr1+sitetr2 + (1|geneticcluster), data=metadata.pup)
summary(model3.sh.1)
```
### Faith's phylogenetic diversity
```{r}
set.seed(123)
model3.pd<-lmer(faith_pd~sitetr1*sitetr2 + (1|geneticcluster), data=metadata.pup)
summary(model3.pd)

set.seed(123)
model3.pd.1<-lmer(faith_pd~sitetr1+sitetr2 + (1|geneticcluster), data=metadata.pup)
summary(model3.pd.1)
```


### Bray-Curtis metric

```{r}
#################################################
#BRAYCURTIS 1000 iterations = results look the same
#################################################
#INTERACTION = NS (0.4787)
#SOM IN TERMS OF SITETR1 = NS (0.3909), IN TERMS OF SITETR2 = almost significant (0.0612)
#SITETR1 = NS = 0.3926
#SITETR2 = almost significant (0.0577)
#################################################
#adonis2
pvalue.adonis2.interaction<-0
pvalue.adonis2.som.sitetr1<-0
pvalue.adonis2.som.sitetr2<-0
pvalue.bray.sitetr1<-0
pvalue.bray.sitetr2<-0
sumofsqs.1<-0
sumofsqs.2<-0
sumofsqs.res<-0
sumofsqs.tot<-0
sumofsqs.sitetr1.1<-0
sumofsqs.sitetr1.res<-0
sumofsqs.sitetr2.1<-0
sumofsqs.sitetr2.res<-0
r2.1<-0
r2.2<-0
r2.res<-0
r2.sitetr1<-0
r2.sitetr2<-0
f.1<-0
f.2<-0
f.sitetr1<-0
f.sitetr2<-0
#betadisper
pvalue.betadisper.sitetr1 <-0
pvalue.betadisper.sitetr2 <-0

n_iterations <- 10
for (i in 1:n_iterations)  
{
  #Take 1 random sample of pup per mom
  meta.pup.1sibling<-metadata.pup %>% group_by(geneticcluster) %>% dplyr::slice_sample(n = 1) %>% as.data.frame()
  #make this subset into a dataframe
  meta_subset_bray<-as.data.frame(meta.pup.1sibling)
  write.table(meta_subset_bray,"meta_subset_bray.txt",sep="\t",row.names=FALSE, quote = FALSE)
  #make a new phyloseq with only 33 pups instead of 72
  psrt.pup.nosibling<-qza_to_phyloseq(
    features="rt_table16s_filtermin10freq_rar30671.qza",
    taxonomy="rt_taxonomysilva.qza",
    metadata = "meta_subset_bray.txt",
    tree = "rt_rooted-tree.qza"
  )
  metadata.psrt.pup.nosibling <- as(sample_data(psrt.pup.nosibling), "data.frame")
  #Calculate distance for this phyloseq
  set.seed(123)
  bray_dist.pup.nosibling = phyloseq::distance(psrt.pup.nosibling, method="bray")
  
  #Calculate adonis for this phyloseq
  set.seed(123)
  pmod.pup.nosibling_adonis2_1<-adonis2(bray_dist.pup.nosibling~sitetr1+sitetr2+sitetr1:sitetr2, data=metadata.psrt.pup.nosibling, by="margin", perm=9999)
  set.seed(123)
  pmod.pup.nosibling_adonis2_2<-adonis2(bray_dist.pup.nosibling~sitetr1+sitetr2, data=metadata.psrt.pup.nosibling, by="margin", perm=9999)
  set.seed(123)
  pmod.pup.nosibling.sitetr1<-adonis2(bray_dist.pup.nosibling~sitetr1, data=metadata.psrt.pup.nosibling, by="terms", perm=9999)
  set.seed(123)
  pmod.pup.nosibling.sitetr2<-adonis2(bray_dist.pup.nosibling~sitetr2, data=metadata.psrt.pup.nosibling, by="terms", perm=9999)
  
  #calculate group dispersion
  set.seed(123)
  betadisper.pup.nosibling.sitetr1 <- betadisper(bray_dist.pup.nosibling, phyloseq::sample_data(psrt.pup.nosibling)$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
  betadisper.pup.nosibling.sitetr1.anova<-anova(betadisper.pup.nosibling.sitetr1)
  set.seed(123)
  betadisper.pup.nosibling.sitetr2 <- betadisper(bray_dist.pup.nosibling, phyloseq::sample_data(psrt.pup.nosibling)$sitetr2, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
  betadisper.pup.nosibling.sitetr2.anova<-anova(betadisper.pup.nosibling.sitetr2)
  
  #summate all the pvalues into one object
  
  #INTERACTION: sitetr1 * sitetr2
  pvalue.adonis2.interaction<-pvalue.adonis2.interaction+pmod.pup.nosibling_adonis2_1$`Pr(>F)`[1]
  #SUM: sitetr1 + sitetr2
  pvalue.adonis2.som.sitetr1<-pvalue.adonis2.som.sitetr1+pmod.pup.nosibling_adonis2_2$`Pr(>F)`[1]
  pvalue.adonis2.som.sitetr2<-pvalue.adonis2.som.sitetr2+pmod.pup.nosibling_adonis2_2$`Pr(>F)`[2]
  sumofsqs.1<-sumofsqs.1+pmod.pup.nosibling_adonis2_2$SumOfSqs[1]
  sumofsqs.2<-sumofsqs.2+pmod.pup.nosibling_adonis2_2$SumOfSqs[2]
  sumofsqs.res<-sumofsqs.res+pmod.pup.nosibling_adonis2_2$SumOfSqs[3]
  r2.1<-r2.1+pmod.pup.nosibling_adonis2_2$R2[1]
  r2.2<-r2.2+pmod.pup.nosibling_adonis2_2$R2[2]
  f.1<-f.1+pmod.pup.nosibling_adonis2_2$F[1]
  f.2<-f.2+pmod.pup.nosibling_adonis2_2$F[2]
  #Site of origin: sitetr1
  pvalue.bray.sitetr1<-pvalue.bray.sitetr1 + pmod.pup.nosibling.sitetr1$`Pr(>F)`[1] 
  sumofsqs.sitetr1.1<-sumofsqs.sitetr1.1+pmod.pup.nosibling.sitetr1$SumOfSqs[1]
  sumofsqs.sitetr1.res<-sumofsqs.sitetr1.res+pmod.pup.nosibling.sitetr1$SumOfSqs[2]
  r2.sitetr1<-r2.sitetr1+pmod.pup.nosibling.sitetr1$R2[1]
  f.sitetr1<-f.sitetr1+pmod.pup.nosibling.sitetr1$F[1]
  #Site of transfer: sitetr2
  pvalue.bray.sitetr2<-pvalue.bray.sitetr2 + pmod.pup.nosibling.sitetr2$`Pr(>F)`[1] 
  sumofsqs.sitetr2.1<-sumofsqs.sitetr2.1+pmod.pup.nosibling.sitetr2$SumOfSqs[1]
  sumofsqs.sitetr2.res<-sumofsqs.sitetr2.res+pmod.pup.nosibling.sitetr2$SumOfSqs[2]
  r2.sitetr2<-r2.sitetr2+pmod.pup.nosibling.sitetr2$R2[1]
  f.sitetr2<-f.sitetr2+pmod.pup.nosibling.sitetr2$F[1]
  #Group dispersion
  pvalue.betadisper.sitetr1<-pvalue.betadisper.sitetr1 + betadisper.pup.nosibling.sitetr1.anova$`Pr(>F)`[1]
  pvalue.betadisper.sitetr2<-pvalue.betadisper.sitetr2 + betadisper.pup.nosibling.sitetr2.anova$`Pr(>F)`[1]
  
}
#RESULTS

#INTERACTION: sitetr1*sitetr2
final_pvalue.bray.interaction<-pvalue.adonis2.interaction/n_iterations #p=0.4787
#SOM: sitetr1 + sitetr2
final_pvalue.adonis2.som.sitetr1 <- pvalue.adonis2.som.sitetr1/n_iterations  #0.3909
final_pvalue.adonis2.som.sitetr2 <- pvalue.adonis2.som.sitetr2/n_iterations  #0.0612
final_ss.som.sitetr1<-sumofsqs.1/n_iterations #
final_ss.som.sitetr2<-sumofsqs.2/n_iterations #
final_ss.som.res<-sumofsqs.res/n_iterations 
calculated_ss.som.total <- final_ss.som.sitetr1+final_ss.som.sitetr2+final_ss.som.res
final_r2.som.sitetr1 <- r2.1/n_iterations #0.03211807
final_r2.som.sitetr2 <- r2.2/n_iterations #0.04200681
calculated_r2.som.res = (1-(final_r2.som.sitetr1)-(final_r2.som.sitetr2))
final_f.1 <- f.1/n_iterations 
final_f.2 <- f.2/n_iterations 

#SITE OF ORIGIN
final_pvalue.bray.sitetr1 <- pvalue.bray.sitetr1/n_iterations
final_sumofsqs.sitetr1.1 <- sumofsqs.sitetr1.1/n_iterations
final_sumofsqs.sitetr1.res <- sumofsqs.sitetr1.res/n_iterations
calculated_ss.sitetr1.total <- final_sumofsqs.sitetr1.1+final_sumofsqs.sitetr1.res
final_r2.sitetr1.1 <- r2.sitetr1/n_iterations
calculated_r2.sitetr1.res = (1-final_r2.sitetr1.1)
final_f.sitetr1 <- f.sitetr1/n_iterations

#SITE OF TRANSFER
final_pvalue.bray.sitetr2 <- pvalue.bray.sitetr2/n_iterations
final_sumofsqs.sitetr2.1 <- sumofsqs.sitetr2.1/n_iterations
final_sumofsqs.sitetr2.res <- sumofsqs.sitetr2.res/n_iterations
calculated_ss.sitetr2.total <- final_sumofsqs.sitetr2.1+final_sumofsqs.sitetr2.res
final_r2.sitetr2.1 <- r2.sitetr2/n_iterations
calculated_r2.sitetr2.res = (1-final_r2.sitetr2.1)
final_f.sitetr2 <- f.sitetr2/n_iterations
#GROUP DISPERSION
final_pvalue.betadisper.sitetr1 <- pvalue.betadisper.sitetr1/n_iterations #0.5822573
final_pvalue.betadisper.sitetr2 <- pvalue.betadisper.sitetr2/n_iterations #0.9971147

#MAKE TABLES
df_som_bray <- data.frame(var=c('sitetr1', 'sitetr2', 'Residual', 'Total'),
                          Df=c(1,1,30,32),
                          SumOfSqs=c(final_ss.som.sitetr1,final_ss.som.sitetr2,final_ss.som.res,calculated_ss.som.total),
                          R2=c(final_r2.som.sitetr1,final_r2.som.sitetr2,calculated_r2.som.res,1),
                          "F"=c(final_f.1, final_f.2 ,"",""),
                          "p"=c(final_pvalue.adonis2.som.sitetr1,final_pvalue.adonis2.som.sitetr2,"",""))

df_sitetr1_bray <- data.frame(var=c('sitetr1', 'Residual', 'Total'),
                              Df=c(1,31,32),
                              SumOfSqs=c(final_sumofsqs.sitetr1.1 , final_sumofsqs.sitetr1.res , calculated_ss.sitetr1.total),
                              R2=c(final_r2.som.sitetr1 , calculated_r2.sitetr1.res , "1"),
                              "F"=c(final_f.sitetr1,"",""),
                              "p"=c(final_pvalue.bray.sitetr1,"",""))

df_sitetr2_bray <- data.frame(var=c('sitetr2', 'Residual', 'Total'),
                              Df=c(1,31,32),
                              SumOfSqs=c(final_sumofsqs.sitetr2.1 , final_sumofsqs.sitetr2.res , calculated_ss.sitetr2.total),
                              R2=c(final_r2.som.sitetr2 , calculated_r2.sitetr2.res , "1"),
                              "F"=c(final_f.sitetr2,"",""),
                              "p"=c(final_pvalue.bray.sitetr2,"",""))

```
**Output table treatment**
```{r}
df_som_bray
write.table(df_som_bray, "adonis2_som_bray.txt")
```
**Output table in terms of site of origin**
```{r}
df_sitetr1_bray
write.table(df_sitetr1_bray, "adonis2_sitetr1_bray.txt")
```
**Betadisper p-value site of origin**
```{r}
final_pvalue.betadisper.sitetr1
```
**Output table in terms of site of transfer**
```{r}
df_sitetr2_bray
write.table(df_sitetr2_bray, "adonis2_sitetr2_bray.txt")
```
**Betadisper p-value site of transfer**
```{r}
final_pvalue.betadisper.sitetr2
```


### Jaccard metric

```{r}
#################################################
#JACCARD 1000 iterations = results look the same
#################################################
#INTERACTION = NS (p=0.5641)
#SOM IN TERMS OF SITETR1 = NS (0.3782) , IN TERMS OF SITETR2 = almost significant(0.0608)
#SITETR1 = NS (0.3794)
#SITETR2 = almost significant (0.0576)
#################################################
#adonis2
pvalue.adonis2.interaction.jacc<-0
pvalue.adonis2.som.sitetr1.jacc<-0
pvalue.adonis2.som.sitetr2.jacc<-0
pvalue.sitetr1.jacc<-0
pvalue.sitetr2.jacc<-0
sumofsqs.1.jacc<-0
sumofsqs.2.jacc<-0
sumofsqs.res.jacc<-0
sumofsqs.tot.jacc<-0
sumofsqs.sitetr1.1.jacc<-0
sumofsqs.sitetr1.res.jacc<-0
sumofsqs.sitetr2.1.jacc<-0
sumofsqs.sitetr2.res.jacc<-0
r2.1.jacc<-0
r2.2.jacc<-0
r2.res.jacc<-0
r2.sitetr1.jacc<-0
r2.sitetr2.jacc<-0
f.1.jacc<-0
f.2.jacc<-0
f.sitetr1.jacc<-0
f.sitetr2.jacc<-0
#betadisper
pvalue.betadisper.sitetr1.jacc <-0
pvalue.betadisper.sitetr2.jacc <-0

n_iterations <- 10
for (i in 1:n_iterations)  
{
  #Take 1 random sample of pup per mom
  meta.pup.1sibling<-metadata.pup %>% group_by(geneticcluster) %>% dplyr::slice_sample(n = 1) %>% as.data.frame()
  #make this subset into a dataframe
  meta_subset_jacc<-as.data.frame(meta.pup.1sibling)
  write.table(meta_subset_jacc,"meta_subset_jacc.txt",sep="\t",row.names=FALSE, quote = FALSE)
  #make a new phyloseq with only 33 pups instead of 72
  psrt.pup.nosibling.jacc<-qza_to_phyloseq(
    features="rt_table16s_filtermin10freq_rar30671.qza",
    taxonomy="rt_taxonomysilva.qza",
    metadata = "meta_subset_jacc.txt",
    tree = "rt_rooted-tree.qza"
  )
  metadata.psrt.pup.nosibling.jacc <- as(sample_data(psrt.pup.nosibling.jacc), "data.frame")
  #Calculate distance for this phyloseq
  set.seed(123)
  jacc_dist.pup.nosibling = phyloseq::distance(psrt.pup.nosibling.jacc, method="jaccard")
  
  #Calculate adonis for this phyloseq
  set.seed(123)
  pmod.pup.nosibling_adonis2_1.jacc <-adonis2(jacc_dist.pup.nosibling~sitetr1+sitetr2+sitetr1:sitetr2, data=metadata.psrt.pup.nosibling.jacc , by="margin", perm=9999)
  set.seed(123)
  pmod.pup.nosibling_adonis2_2.jacc <-adonis2(jacc_dist.pup.nosibling~sitetr1+sitetr2, data=metadata.psrt.pup.nosibling.jacc , by="margin", perm=9999)
  set.seed(123)
  pmod.pup.nosibling.sitetr1.jacc <-adonis2(jacc_dist.pup.nosibling~sitetr1, data=metadata.psrt.pup.nosibling.jacc , by="terms", perm=9999)
  set.seed(123)
  pmod.pup.nosibling.sitetr2.jacc <-adonis2(jacc_dist.pup.nosibling~sitetr2, data=metadata.psrt.pup.nosibling.jacc , by="terms", perm=9999)

  #calculate group dispersion
  set.seed(123)
  betadisper.pup.nosibling.sitetr1.jacc  <- betadisper(jacc_dist.pup.nosibling, phyloseq::sample_data(psrt.pup.nosibling.jacc)$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
  betadisper.pup.nosibling.sitetr1.anova.jacc <-anova(betadisper.pup.nosibling.sitetr1.jacc)
  set.seed(123)
  betadisper.pup.nosibling.sitetr2.jacc  <- betadisper(jacc_dist.pup.nosibling, phyloseq::sample_data(psrt.pup.nosibling.jacc)$sitetr2, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
  betadisper.pup.nosibling.sitetr2.anova.jacc<-anova(betadisper.pup.nosibling.sitetr2.jacc)
  
  #summate all the pvalues into one object

  #INTERACTION: sitetr1 * sitetr2
  pvalue.adonis2.interaction.jacc<-pvalue.adonis2.interaction.jacc+pmod.pup.nosibling_adonis2_1.jacc$`Pr(>F)`[1]
  #SUM: sitetr1 + sitetr2
  pvalue.adonis2.som.sitetr1.jacc<-pvalue.adonis2.som.sitetr1.jacc+pmod.pup.nosibling_adonis2_2.jacc$`Pr(>F)`[1]
  pvalue.adonis2.som.sitetr2.jacc<-pvalue.adonis2.som.sitetr2.jacc+pmod.pup.nosibling_adonis2_2.jacc$`Pr(>F)`[2]
  sumofsqs.1.jacc<-sumofsqs.1.jacc+pmod.pup.nosibling_adonis2_2.jacc$SumOfSqs[1]
  sumofsqs.2.jacc<-sumofsqs.2.jacc+pmod.pup.nosibling_adonis2_2.jacc$SumOfSqs[2]
  sumofsqs.res.jacc<-sumofsqs.res.jacc+pmod.pup.nosibling_adonis2_2.jacc$SumOfSqs[3]
  r2.1.jacc<-r2.1.jacc+pmod.pup.nosibling_adonis2_2.jacc$R2[1]
  r2.2.jacc<-r2.2.jacc+pmod.pup.nosibling_adonis2_2.jacc$R2[2]
  f.1.jacc<-f.1.jacc+pmod.pup.nosibling_adonis2_2.jacc$F[1]
  f.2.jacc<-f.2.jacc+pmod.pup.nosibling_adonis2_2.jacc$F[2]
  #Site of origin: sitetr1
  pvalue.sitetr1.jacc<-pvalue.sitetr1.jacc + pmod.pup.nosibling.sitetr1.jacc$`Pr(>F)`[1] 
  sumofsqs.sitetr1.1.jacc<-sumofsqs.sitetr1.1.jacc+pmod.pup.nosibling.sitetr1.jacc$SumOfSqs[1]
  sumofsqs.sitetr1.res.jacc<-sumofsqs.sitetr1.res.jacc+pmod.pup.nosibling.sitetr1.jacc$SumOfSqs[2]
  r2.sitetr1.jacc<-r2.sitetr1.jacc+pmod.pup.nosibling.sitetr1.jacc$R2[1]
  f.sitetr1.jacc<-f.sitetr1.jacc+pmod.pup.nosibling.sitetr1.jacc$F[1]
  #Site of transfer: sitetr2
  pvalue.sitetr2.jacc<-pvalue.sitetr2.jacc + pmod.pup.nosibling.sitetr2.jacc$`Pr(>F)`[1] 
  sumofsqs.sitetr2.1.jacc<-sumofsqs.sitetr2.1.jacc+pmod.pup.nosibling.sitetr2.jacc$SumOfSqs[1]
  sumofsqs.sitetr2.res.jacc<-sumofsqs.sitetr2.res.jacc+pmod.pup.nosibling.sitetr2.jacc$SumOfSqs[2]
  r2.sitetr2.jacc<-r2.sitetr2.jacc+pmod.pup.nosibling.sitetr2.jacc$R2[1]
  f.sitetr2.jacc<-f.sitetr2.jacc+pmod.pup.nosibling.sitetr2.jacc$F[1]
  #Group dispersion
  pvalue.betadisper.sitetr1.jacc<-pvalue.betadisper.sitetr1.jacc + betadisper.pup.nosibling.sitetr1.anova.jacc$`Pr(>F)`[1]
  pvalue.betadisper.sitetr2.jacc<-pvalue.betadisper.sitetr2.jacc + betadisper.pup.nosibling.sitetr2.anova.jacc$`Pr(>F)`[1]
  
}
#RESULTS

#INTERACTION: sitetr1*sitetr2
final_pvalue.interaction.jacc<-pvalue.adonis2.interaction.jacc/n_iterations #p=0.5641
#SOM: sitetr1 + sitetr2
final_pvalue.adonis2.som.sitetr1.jacc <- pvalue.adonis2.som.sitetr1.jacc/n_iterations  #0.3782
final_pvalue.adonis2.som.sitetr2.jacc <- pvalue.adonis2.som.sitetr2.jacc/n_iterations  #0.0608
final_ss.som.sitetr1.jacc<-sumofsqs.1.jacc/n_iterations #
final_ss.som.sitetr2.jacc<-sumofsqs.2.jacc/n_iterations #
final_ss.som.res.jacc<-sumofsqs.res.jacc/n_iterations 
calculated_ss.som.total.jacc <- final_ss.som.sitetr1.jacc+final_ss.som.sitetr2.jacc+final_ss.som.res.jacc
final_r2.som.sitetr1.jacc <- r2.1.jacc/n_iterations #0.
final_r2.som.sitetr2.jacc <- r2.2.jacc/n_iterations #0.
calculated_r2.som.res.jacc = (1-(final_r2.som.sitetr1.jacc)-(final_r2.som.sitetr2.jacc))
final_f.1.jacc <- f.1.jacc/n_iterations 
final_f.2.jacc <- f.2.jacc/n_iterations 

#SITE OF ORIGIN
final_pvalue.sitetr1.jacc <- pvalue.sitetr1.jacc/n_iterations #0.3794
final_sumofsqs.sitetr1.1.jacc <- sumofsqs.sitetr1.1.jacc/n_iterations
final_sumofsqs.sitetr1.res.jacc <- sumofsqs.sitetr1.res.jacc/n_iterations
calculated_ss.sitetr1.total.jacc <- final_sumofsqs.sitetr1.1.jacc+final_sumofsqs.sitetr1.res.jacc
final_r2.sitetr1.1.jacc <- r2.sitetr1.jacc/n_iterations
calculated_r2.sitetr1.res.jacc = (1-final_r2.sitetr1.1.jacc)
final_f.sitetr1.jacc <- f.sitetr1.jacc/n_iterations

#SITE OF TRANSFER
final_pvalue.sitetr2.jacc <- pvalue.sitetr2.jacc/n_iterations #0.0576
final_sumofsqs.sitetr2.1.jacc <- sumofsqs.sitetr2.1.jacc/n_iterations
final_sumofsqs.sitetr2.res.jacc <- sumofsqs.sitetr2.res.jacc/n_iterations
calculated_ss.sitetr2.total.jacc <- final_sumofsqs.sitetr2.1.jacc+final_sumofsqs.sitetr2.res.jacc
final_r2.sitetr2.1.jacc <- r2.sitetr2.jacc/n_iterations
calculated_r2.sitetr2.res.jacc = (1-final_r2.sitetr2.1.jacc)
final_f.sitetr2.jacc<- f.sitetr2.jacc/n_iterations

#GROUP DISPERSION
final_pvalue.betadisper.sitetr1.jacc <- pvalue.betadisper.sitetr1.jacc/n_iterations #0.6447998
final_pvalue.betadisper.sitetr2.jacc <- pvalue.betadisper.sitetr2.jacc/n_iterations #0.9854238

#MAKE TABLES
df_som_jaccard <- data.frame(var=c('sitetr1', 'sitetr2', 'Residual', 'Total'),
                          Df=c(1,1,30,32),
                          SumOfSqs=c(final_ss.som.sitetr1.jacc,final_ss.som.sitetr2.jacc,final_ss.som.res.jacc,calculated_ss.som.total.jacc),
                          R2=c(final_r2.som.sitetr1.jacc,final_r2.som.sitetr2.jacc,calculated_r2.som.res.jacc,1),
                          "F"=c(final_f.1, final_f.2.jacc ,"",""),
                          "p"=c(final_pvalue.adonis2.som.sitetr1.jacc,final_pvalue.adonis2.som.sitetr2.jacc,"",""))

df_sitetr1_jaccard <- data.frame(var=c('sitetr1', 'Residual', 'Total'),
                              Df=c(1,31,32),
                              SumOfSqs=c(final_sumofsqs.sitetr1.1.jacc , final_sumofsqs.sitetr1.res.jacc , calculated_ss.sitetr1.total.jacc),
                              R2=c(final_r2.som.sitetr1.jacc , calculated_r2.sitetr1.res.jacc , "1"),
                              "F"=c(final_f.sitetr1.jacc,"",""),
                              "p"=c(final_pvalue.sitetr1.jacc,"",""))

df_sitetr2_jaccard <- data.frame(var=c('sitetr2', 'Residual', 'Total'),
                              Df=c(1,31,32),
                              SumOfSqs=c(final_sumofsqs.sitetr2.1.jacc , final_sumofsqs.sitetr2.res.jacc , calculated_ss.sitetr2.total.jacc),
                              R2=c(final_r2.som.sitetr2.jacc , calculated_r2.sitetr2.res.jacc , "1"),
                              "F"=c(final_f.sitetr2.jacc,"",""),
                              "p"=c(final_pvalue.sitetr2.jacc,"",""))

#################################################
```
**Output table treatment**
```{r}
df_som_jaccard
write.table(df_som_jaccard, "adonis2_som_jaccard.txt")
```
**Output table in terms of site of origin**
```{r}
df_sitetr1_jaccard
write.table(df_sitetr1_jaccard, "adonis2_sitetr1_jaccard.txt")
```
**Betadisper p-value site of origin**
```{r}
final_pvalue.betadisper.sitetr1.jacc
```
**Output table in terms of site of transfer**
```{r}
df_sitetr2_jaccard
write.table(df_sitetr2_jaccard, "adonis2_sitetr2_jaccard.txt")
```
**Betadisper p-value site of transfer**
```{r}
final_pvalue.betadisper.sitetr2.jacc
```

### Weighted Unifrac metric

```{r}
#################################################
#WEIGHTED UNIFRAC 1000 iterations = results look the same
#################################################
#INTERACTION = NS (0.1747)
#SOM IN TERMS OF SITETR1 = NS (0.2498), IN TERMS OF SITETR2 = NS (0.4087)
#SITETR1 = NS (0.2193)
#SITETR2 = NS (0.3679)
#################################################
#adonis2
pvalue.adonis2.interaction.wunifrac<-0
pvalue.adonis2.som.sitetr1.wunifrac<-0
pvalue.adonis2.som.sitetr2.wunifrac<-0
pvalue.sitetr1.wunifrac<-0
pvalue.sitetr2.wunifrac<-0
sumofsqs.1.wunifrac<-0
sumofsqs.2.wunifrac<-0
sumofsqs.res.wunifrac<-0
sumofsqs.tot.wunifrac<-0
sumofsqs.sitetr1.1.wunifrac<-0
sumofsqs.sitetr1.res.wunifrac<-0
sumofsqs.sitetr2.1.wunifrac<-0
sumofsqs.sitetr2.res.wunifrac<-0
r2.1.wunifrac<-0
r2.2.wunifrac<-0
r2.res.wunifrac<-0
r2.sitetr1.wunifrac<-0
r2.sitetr2.wunifrac<-0
f.1.wunifrac<-0
f.2.wunifrac<-0
f.sitetr1.wunifrac<-0
f.sitetr2.wunifrac<-0
#betadisper
pvalue.betadisper.sitetr1.wunifrac <-0
pvalue.betadisper.sitetr2.wunifrac <-0

n_iterations <- 10
for (i in 1:n_iterations)  
{
  #Take 1 random sample of pup per mom
  meta.pup.1sibling<-metadata.pup %>% group_by(geneticcluster) %>% dplyr::slice_sample(n = 1) %>% as.data.frame()
  #make this subset into a dataframe
  meta_subset_wunifrac<-as.data.frame(meta.pup.1sibling)
  write.table(meta_subset_wunifrac,"meta_subset_wunifrac.txt",sep="\t",row.names=FALSE, quote = FALSE)
  #make a new phyloseq with only 33 pups instead of 72
  psrt.pup.nosibling.wunifrac<-qza_to_phyloseq(
    features="rt_table16s_filtermin10freq_rar30671.qza",
    taxonomy="rt_taxonomysilva.qza",
    metadata = "meta_subset_wunifrac.txt",
    tree = "rt_rooted-tree.qza"
  )
  metadata.psrt.pup.nosibling.wunifrac <- as(sample_data(psrt.pup.nosibling.wunifrac), "data.frame")
  #Calculate distance for this phyloseq
  set.seed(123)
  wunifrac_dist.pup.nosibling = phyloseq::distance(psrt.pup.nosibling.wunifrac, method="wunifrac")
  
  #Calculate adonis for this phyloseq
  set.seed(123)
  pmod.pup.nosibling_adonis2_1.wunifrac <-adonis2(wunifrac_dist.pup.nosibling~sitetr1+sitetr2+sitetr1:sitetr2, data=metadata.psrt.pup.nosibling.wunifrac , by="margin", perm=9999)
  set.seed(123)
  pmod.pup.nosibling_adonis2_2.wunifrac <-adonis2(wunifrac_dist.pup.nosibling~sitetr1+sitetr2, data=metadata.psrt.pup.nosibling.wunifrac , by="margin", perm=9999)
  set.seed(123)
  pmod.pup.nosibling.sitetr1.wunifrac <-adonis2(wunifrac_dist.pup.nosibling~sitetr1, data=metadata.psrt.pup.nosibling.wunifrac , by="terms", perm=9999)
  set.seed(123)
  pmod.pup.nosibling.sitetr2.wunifrac <-adonis2(wunifrac_dist.pup.nosibling~sitetr2, data=metadata.psrt.pup.nosibling.wunifrac , by="terms", perm=9999)
  
  #calculate group dispersion
  set.seed(123)
  betadisper.pup.nosibling.sitetr1.wunifrac  <- betadisper(wunifrac_dist.pup.nosibling, phyloseq::sample_data(psrt.pup.nosibling.wunifrac)$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
  betadisper.pup.nosibling.sitetr1.anova.wunifrac <-anova(betadisper.pup.nosibling.sitetr1.wunifrac)
  set.seed(123)
  betadisper.pup.nosibling.sitetr2.wunifrac  <- betadisper(wunifrac_dist.pup.nosibling, phyloseq::sample_data(psrt.pup.nosibling.wunifrac)$sitetr2, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
  betadisper.pup.nosibling.sitetr2.anova.wunifrac<-anova(betadisper.pup.nosibling.sitetr2.wunifrac)
  
  #summate all the pvalues into one object
  
  #INTERACTION: sitetr1 * sitetr2
  pvalue.adonis2.interaction.wunifrac<-pvalue.adonis2.interaction.wunifrac+pmod.pup.nosibling_adonis2_1.wunifrac$`Pr(>F)`[1]
  #SUM: sitetr1 + sitetr2
  pvalue.adonis2.som.sitetr1.wunifrac<-pvalue.adonis2.som.sitetr1.wunifrac+pmod.pup.nosibling_adonis2_2.wunifrac$`Pr(>F)`[1]
  pvalue.adonis2.som.sitetr2.wunifrac<-pvalue.adonis2.som.sitetr2.wunifrac+pmod.pup.nosibling_adonis2_2.wunifrac$`Pr(>F)`[2]
  sumofsqs.1.wunifrac<-sumofsqs.1.wunifrac+pmod.pup.nosibling_adonis2_2.wunifrac$SumOfSqs[1]
  sumofsqs.2.wunifrac<-sumofsqs.2.wunifrac+pmod.pup.nosibling_adonis2_2.wunifrac$SumOfSqs[2]
  sumofsqs.res.wunifrac<-sumofsqs.res.wunifrac+pmod.pup.nosibling_adonis2_2.wunifrac$SumOfSqs[3]
  r2.1.wunifrac<-r2.1.wunifrac+pmod.pup.nosibling_adonis2_2.wunifrac$R2[1]
  r2.2.wunifrac<-r2.2.wunifrac+pmod.pup.nosibling_adonis2_2.wunifrac$R2[2]
  f.1.wunifrac<-f.1.wunifrac+pmod.pup.nosibling_adonis2_2.wunifrac$F[1]
  f.2.wunifrac<-f.2.wunifrac+pmod.pup.nosibling_adonis2_2.wunifrac$F[2]
  #Site of origin: sitetr1
  pvalue.sitetr1.wunifrac<-pvalue.sitetr1.wunifrac + pmod.pup.nosibling.sitetr1.wunifrac$`Pr(>F)`[1] 
  sumofsqs.sitetr1.1.wunifrac<-sumofsqs.sitetr1.1.wunifrac+pmod.pup.nosibling.sitetr1.wunifrac$SumOfSqs[1]
  sumofsqs.sitetr1.res.wunifrac<-sumofsqs.sitetr1.res.wunifrac+pmod.pup.nosibling.sitetr1.wunifrac$SumOfSqs[2]
  r2.sitetr1.wunifrac<-r2.sitetr1.wunifrac+pmod.pup.nosibling.sitetr1.wunifrac$R2[1]
  f.sitetr1.wunifrac<-f.sitetr1.wunifrac+pmod.pup.nosibling.sitetr1.wunifrac$F[1]
  #Site of transfer: sitetr2
  pvalue.sitetr2.wunifrac<-pvalue.sitetr2.wunifrac + pmod.pup.nosibling.sitetr2.wunifrac$`Pr(>F)`[1] 
  sumofsqs.sitetr2.1.wunifrac<-sumofsqs.sitetr2.1.wunifrac+pmod.pup.nosibling.sitetr2.wunifrac$SumOfSqs[1]
  sumofsqs.sitetr2.res.wunifrac<-sumofsqs.sitetr2.res.wunifrac+pmod.pup.nosibling.sitetr2.wunifrac$SumOfSqs[2]
  r2.sitetr2.wunifrac<-r2.sitetr2.wunifrac+pmod.pup.nosibling.sitetr2.wunifrac$R2[1]
  f.sitetr2.wunifrac<-f.sitetr2.wunifrac+pmod.pup.nosibling.sitetr2.wunifrac$F[1]
  #Group dispersion
  pvalue.betadisper.sitetr1.wunifrac<-pvalue.betadisper.sitetr1.wunifrac + betadisper.pup.nosibling.sitetr1.anova.wunifrac$`Pr(>F)`[1]
  pvalue.betadisper.sitetr2.wunifrac<-pvalue.betadisper.sitetr2.wunifrac + betadisper.pup.nosibling.sitetr2.anova.wunifrac$`Pr(>F)`[1]
  
}
#RESULTS

#INTERACTION: sitetr1*sitetr2
final_pvalue.interaction.wunifrac<-pvalue.adonis2.interaction.wunifrac/n_iterations #p=0.1747
#SOM: sitetr1 + sitetr2
final_pvalue.adonis2.som.sitetr1.wunifrac <- pvalue.adonis2.som.sitetr1.wunifrac/n_iterations  #0.2498
final_pvalue.adonis2.som.sitetr2.wunifrac <- pvalue.adonis2.som.sitetr2.wunifrac/n_iterations  #0.4087
final_ss.som.sitetr1.wunifrac<-sumofsqs.1.wunifrac/n_iterations #
final_ss.som.sitetr2.wunifrac<-sumofsqs.2.wunifrac/n_iterations #
final_ss.som.res.wunifrac<-sumofsqs.res.wunifrac/n_iterations 
calculated_ss.som.total.wunifrac <- final_ss.som.sitetr1.wunifrac+final_ss.som.sitetr2.wunifrac+final_ss.som.res.wunifrac
final_r2.som.sitetr1.wunifrac <- r2.1.wunifrac/n_iterations #0.
final_r2.som.sitetr2.wunifrac <- r2.2.wunifrac/n_iterations #0.
calculated_r2.som.res.wunifrac = (1-(final_r2.som.sitetr1.wunifrac)-(final_r2.som.sitetr2.wunifrac))
final_f.1.wunifrac <- f.1.wunifrac/n_iterations 
final_f.2.wunifrac <- f.2.wunifrac/n_iterations 

#SITE OF ORIGIN
final_pvalue.sitetr1.wunifrac <- pvalue.sitetr1.wunifrac/n_iterations
final_sumofsqs.sitetr1.1.wunifrac <- sumofsqs.sitetr1.1.wunifrac/n_iterations
final_sumofsqs.sitetr1.res.wunifrac <- sumofsqs.sitetr1.res.wunifrac/n_iterations
calculated_ss.sitetr1.total.wunifrac <- final_sumofsqs.sitetr1.1.wunifrac+final_sumofsqs.sitetr1.res.wunifrac
final_r2.sitetr1.1.wunifrac <- r2.sitetr1.wunifrac/n_iterations
calculated_r2.sitetr1.res.wunifrac = (1-final_r2.sitetr1.1.wunifrac)
final_f.sitetr1.wunifrac <- f.sitetr1.wunifrac/n_iterations

#SITE OF TRANSFER
final_pvalue.sitetr2.wunifrac <- pvalue.sitetr2.wunifrac/n_iterations
final_sumofsqs.sitetr2.1.wunifrac <- sumofsqs.sitetr2.1.wunifrac/n_iterations
final_sumofsqs.sitetr2.res.wunifrac <- sumofsqs.sitetr2.res.wunifrac/n_iterations
calculated_ss.sitetr2.total.wunifrac <- final_sumofsqs.sitetr2.1.wunifrac+final_sumofsqs.sitetr2.res.wunifrac
final_r2.sitetr2.1.wunifrac <- r2.sitetr2.wunifrac/n_iterations
calculated_r2.sitetr2.res.wunifrac = (1-final_r2.sitetr2.1.wunifrac)
final_f.sitetr2.wunifrac<- f.sitetr2.wunifrac/n_iterations

#GROUP DISPERSION
final_pvalue.betadisper.sitetr1.wunifrac <- pvalue.betadisper.sitetr1.wunifrac/n_iterations #0.8
final_pvalue.betadisper.sitetr2.wunifrac <- pvalue.betadisper.sitetr2.wunifrac/n_iterations #0.48

#MAKE TABLES
df_som_wunifrac <- data.frame(var=c('sitetr1', 'sitetr2', 'Residual', 'Total'),
                                 Df=c(1,1,30,32),
                                 SumOfSqs=c(final_ss.som.sitetr1.wunifrac,final_ss.som.sitetr2.wunifrac,final_ss.som.res.wunifrac,calculated_ss.som.total.wunifrac),
                                 R2=c(final_r2.som.sitetr1.wunifrac,final_r2.som.sitetr2.wunifrac,calculated_r2.som.res.wunifrac,1),
                                 "F"=c(final_f.1, final_f.2.wunifrac ,"",""),
                                 "p"=c(final_pvalue.adonis2.som.sitetr1.wunifrac,final_pvalue.adonis2.som.sitetr2.wunifrac,"",""))

df_sitetr1_wunifrac <- data.frame(var=c('sitetr1', 'Residual', 'Total'),
                                     Df=c(1,31,32),
                                     SumOfSqs=c(final_sumofsqs.sitetr1.1.wunifrac , final_sumofsqs.sitetr1.res.wunifrac , calculated_ss.sitetr1.total.wunifrac),
                                     R2=c(final_r2.som.sitetr1.wunifrac , calculated_r2.sitetr1.res.wunifrac , "1"),
                                     "F"=c(final_f.sitetr1.wunifrac,"",""),
                                     "p"=c(final_pvalue.sitetr1.wunifrac,"",""))

df_sitetr2_wunifrac <- data.frame(var=c('sitetr2', 'Residual', 'Total'),
                                     Df=c(1,31,32),
                                     SumOfSqs=c(final_sumofsqs.sitetr2.1.wunifrac , final_sumofsqs.sitetr2.res.wunifrac , calculated_ss.sitetr2.total.wunifrac),
                                     R2=c(final_r2.som.sitetr2.wunifrac , calculated_r2.sitetr2.res.wunifrac , "1"),
                                     "F"=c(final_f.sitetr2.wunifrac,"",""),
                                     "p"=c(final_pvalue.sitetr2.wunifrac,"",""))

#################################################
```
**Output table treatment**
```{r}
df_som_wunifrac
write.table(df_som_wunifrac, "adonis2_som_wunifrac.txt")
```
**Output table in terms of site of origin**
```{r}
df_sitetr1_wunifrac
write.table(df_sitetr1_wunifrac, "adonis2_sitetr1_wunifrac.txt")
```
**Betadisper p-value site of origin**
```{r}
final_pvalue.betadisper.sitetr1.wunifrac
```
**Output table in terms of site of transfer**
```{r}
df_sitetr2_wunifrac
write.table(df_sitetr2_wunifrac, "adonis2_sitetr2_wunifrac.txt")
```
**Betadisper p-value site of transfer**
```{r}
final_pvalue.betadisper.sitetr2.wunifrac
```

### Unweighted Unifrac metric

```{r}
#################################################
#UNWEIGHTED UNIFRAC 1000 iterations = results look the same
#################################################
#INTERACTION = NS (0.534)
#SOM IN TERMS OF SITETR1 = almost significant (p=0.0845), IN TERMS OF SITETR2 = almost significant (p=0.1037)
#SITETR1 = almost significant (p=0.0876)
#SITETR2 = almost significant (p=0.1073)
#################################################
#################################################
#adonis 2
pvalue.adonis2.interaction.unwunifrac<-0
pvalue.adonis2.som.sitetr1.unwunifrac<-0
pvalue.adonis2.som.sitetr2.unwunifrac<-0
pvalue.sitetr1.unwunifrac<-0
pvalue.sitetr2.unwunifrac<-0
sumofsqs.1.unwunifrac<-0
sumofsqs.2.unwunifrac<-0
sumofsqs.res.unwunifrac<-0
sumofsqs.tot.unwunifrac<-0
sumofsqs.sitetr1.1.unwunifrac<-0
sumofsqs.sitetr1.res.unwunifrac<-0
sumofsqs.sitetr2.1.unwunifrac<-0
sumofsqs.sitetr2.res.unwunifrac<-0
r2.1.unwunifrac<-0
r2.2.unwunifrac<-0
r2.res.unwunifrac<-0
r2.sitetr1.unwunifrac<-0
r2.sitetr2.unwunifrac<-0
f.1.unwunifrac<-0
f.2.unwunifrac<-0
f.sitetr1.unwunifrac<-0
f.sitetr2.unwunifrac<-0
#betadisper
pvalue.betadisper.sitetr1.unwunifrac <-0
pvalue.betadisper.sitetr2.unwunifrac <-0

n_iterations <- 10
for (i in 1:n_iterations)  
{
  #Take 1 random sample of pup per mom
  meta.pup.1sibling<-metadata.pup %>% group_by(geneticcluster) %>% dplyr::slice_sample(n = 1) %>% as.data.frame()
  #make this subset into a dataframe
  meta_subset_unwunifrac<-as.data.frame(meta.pup.1sibling)
  write.table(meta_subset_unwunifrac,"meta_subset_unwunifrac.txt",sep="\t",row.names=FALSE, quote = FALSE)
  #make a new phyloseq with only 33 pups instead of 72
  psrt.pup.nosibling.unwunifrac<-qza_to_phyloseq(
    features="rt_table16s_filtermin10freq_rar30671.qza",
    taxonomy="rt_taxonomysilva.qza",
    metadata = "meta_subset_unwunifrac.txt",
    tree = "rt_rooted-tree.qza"
  )
  metadata.psrt.pup.nosibling.unwunifrac <- as(sample_data(psrt.pup.nosibling.unwunifrac), "data.frame")
  #Calculate distance for this phyloseq
  set.seed(123)
  unwunifrac_dist.pup.nosibling = phyloseq::distance(psrt.pup.nosibling.unwunifrac, method="unwunifrac")
  
  #Calculate adonis for this phyloseq
  set.seed(123)
  pmod.pup.nosibling_adonis2_1.unwunifrac <-adonis2(unwunifrac_dist.pup.nosibling~sitetr1+sitetr2+sitetr1:sitetr2, data=metadata.psrt.pup.nosibling.unwunifrac , by="margin", perm=9999)
  set.seed(123)
  pmod.pup.nosibling_adonis2_2.unwunifrac <-adonis2(unwunifrac_dist.pup.nosibling~sitetr1+sitetr2, data=metadata.psrt.pup.nosibling.unwunifrac , by="margin", perm=9999)
  set.seed(123)
  pmod.pup.nosibling.sitetr1.unwunifrac <-adonis2(unwunifrac_dist.pup.nosibling~sitetr1, data=metadata.psrt.pup.nosibling.unwunifrac , by="terms", perm=9999)
  set.seed(123)
  pmod.pup.nosibling.sitetr2.unwunifrac <-adonis2(unwunifrac_dist.pup.nosibling~sitetr2, data=metadata.psrt.pup.nosibling.unwunifrac , by="terms", perm=9999)
  
  #calculate group dispersion
  set.seed(123)
  betadisper.pup.nosibling.sitetr1.unwunifrac  <- betadisper(unwunifrac_dist.pup.nosibling, phyloseq::sample_data(psrt.pup.nosibling.unwunifrac)$sitetr1, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
  betadisper.pup.nosibling.sitetr1.anova.unwunifrac <-anova(betadisper.pup.nosibling.sitetr1.unwunifrac)
  set.seed(123)
  betadisper.pup.nosibling.sitetr2.unwunifrac  <- betadisper(unwunifrac_dist.pup.nosibling, phyloseq::sample_data(psrt.pup.nosibling.unwunifrac)$sitetr2, bias.adjust = TRUE, type=c("centroid"), sqrt.dist = FALSE, add = FALSE)
  betadisper.pup.nosibling.sitetr2.anova.unwunifrac<-anova(betadisper.pup.nosibling.sitetr2.unwunifrac)
  
  #summate all the pvalues into one object
  
  #INTERACTION: sitetr1 * sitetr2
  pvalue.adonis2.interaction.unwunifrac<-pvalue.adonis2.interaction.unwunifrac+pmod.pup.nosibling_adonis2_1.unwunifrac$`Pr(>F)`[1]
  #SUM: sitetr1 + sitetr2
  pvalue.adonis2.som.sitetr1.unwunifrac<-pvalue.adonis2.som.sitetr1.unwunifrac+pmod.pup.nosibling_adonis2_2.unwunifrac$`Pr(>F)`[1]
  pvalue.adonis2.som.sitetr2.unwunifrac<-pvalue.adonis2.som.sitetr2.unwunifrac+pmod.pup.nosibling_adonis2_2.unwunifrac$`Pr(>F)`[2]
  sumofsqs.1.unwunifrac<-sumofsqs.1.unwunifrac+pmod.pup.nosibling_adonis2_2.unwunifrac$SumOfSqs[1]
  sumofsqs.2.unwunifrac<-sumofsqs.2.unwunifrac+pmod.pup.nosibling_adonis2_2.unwunifrac$SumOfSqs[2]
  sumofsqs.res.unwunifrac<-sumofsqs.res.unwunifrac+pmod.pup.nosibling_adonis2_2.unwunifrac$SumOfSqs[3]
  r2.1.unwunifrac<-r2.1.unwunifrac+pmod.pup.nosibling_adonis2_2.unwunifrac$R2[1]
  r2.2.unwunifrac<-r2.2.unwunifrac+pmod.pup.nosibling_adonis2_2.unwunifrac$R2[2]
  f.1.unwunifrac<-f.1.unwunifrac+pmod.pup.nosibling_adonis2_2.unwunifrac$F[1]
  f.2.unwunifrac<-f.2.unwunifrac+pmod.pup.nosibling_adonis2_2.unwunifrac$F[2]
  #Site of origin: sitetr1
  pvalue.sitetr1.unwunifrac<-pvalue.sitetr1.unwunifrac + pmod.pup.nosibling.sitetr1.unwunifrac$`Pr(>F)`[1] 
  sumofsqs.sitetr1.1.unwunifrac<-sumofsqs.sitetr1.1.unwunifrac+pmod.pup.nosibling.sitetr1.unwunifrac$SumOfSqs[1]
  sumofsqs.sitetr1.res.unwunifrac<-sumofsqs.sitetr1.res.unwunifrac+pmod.pup.nosibling.sitetr1.unwunifrac$SumOfSqs[2]
  r2.sitetr1.unwunifrac<-r2.sitetr1.unwunifrac+pmod.pup.nosibling.sitetr1.unwunifrac$R2[1]
  f.sitetr1.unwunifrac<-f.sitetr1.unwunifrac+pmod.pup.nosibling.sitetr1.unwunifrac$F[1]
  #Site of transfer: sitetr2
  pvalue.sitetr2.unwunifrac<-pvalue.sitetr2.unwunifrac + pmod.pup.nosibling.sitetr2.unwunifrac$`Pr(>F)`[1] 
  sumofsqs.sitetr2.1.unwunifrac<-sumofsqs.sitetr2.1.unwunifrac+pmod.pup.nosibling.sitetr2.unwunifrac$SumOfSqs[1]
  sumofsqs.sitetr2.res.unwunifrac<-sumofsqs.sitetr2.res.unwunifrac+pmod.pup.nosibling.sitetr2.unwunifrac$SumOfSqs[2]
  r2.sitetr2.unwunifrac<-r2.sitetr2.unwunifrac+pmod.pup.nosibling.sitetr2.unwunifrac$R2[1]
  f.sitetr2.unwunifrac<-f.sitetr2.unwunifrac+pmod.pup.nosibling.sitetr2.unwunifrac$F[1]
  #Group dispersion
  pvalue.betadisper.sitetr1.unwunifrac<-pvalue.betadisper.sitetr1.unwunifrac + betadisper.pup.nosibling.sitetr1.anova.unwunifrac$`Pr(>F)`[1]
  pvalue.betadisper.sitetr2.unwunifrac<-pvalue.betadisper.sitetr2.unwunifrac + betadisper.pup.nosibling.sitetr2.anova.unwunifrac$`Pr(>F)`[1]
  
}
#RESULTS

#INTERACTION: sitetr1*sitetr2
final_pvalue.interaction.unwunifrac<-pvalue.adonis2.interaction.unwunifrac/n_iterations #p=0.534
#SOM: sitetr1 + sitetr2
final_pvalue.adonis2.som.sitetr1.unwunifrac <- pvalue.adonis2.som.sitetr1.unwunifrac/n_iterations  #0.0845
final_pvalue.adonis2.som.sitetr2.unwunifrac <- pvalue.adonis2.som.sitetr2.unwunifrac/n_iterations  #0.1037
final_ss.som.sitetr1.unwunifrac<-sumofsqs.1.unwunifrac/n_iterations #
final_ss.som.sitetr2.unwunifrac<-sumofsqs.2.unwunifrac/n_iterations #
final_ss.som.res.unwunifrac<-sumofsqs.res.unwunifrac/n_iterations 
calculated_ss.som.total.unwunifrac <- final_ss.som.sitetr1.unwunifrac+final_ss.som.sitetr2.unwunifrac+final_ss.som.res.unwunifrac
final_r2.som.sitetr1.unwunifrac <- r2.1.unwunifrac/n_iterations #0.
final_r2.som.sitetr2.unwunifrac <- r2.2.unwunifrac/n_iterations #0.
calculated_r2.som.res.unwunifrac = (1-(final_r2.som.sitetr1.unwunifrac)-(final_r2.som.sitetr2.unwunifrac))
final_f.1.unwunifrac <- f.1.unwunifrac/n_iterations 
final_f.2.unwunifrac <- f.2.unwunifrac/n_iterations 

#SITE OF ORIGIN
final_pvalue.sitetr1.unwunifrac <- pvalue.sitetr1.unwunifrac/n_iterations
final_sumofsqs.sitetr1.1.unwunifrac <- sumofsqs.sitetr1.1.unwunifrac/n_iterations
final_sumofsqs.sitetr1.res.unwunifrac <- sumofsqs.sitetr1.res.unwunifrac/n_iterations
calculated_ss.sitetr1.total.unwunifrac <- final_sumofsqs.sitetr1.1.unwunifrac+final_sumofsqs.sitetr1.res.unwunifrac
final_r2.sitetr1.1.unwunifrac <- r2.sitetr1.unwunifrac/n_iterations
calculated_r2.sitetr1.res.unwunifrac = (1-final_r2.sitetr1.1.unwunifrac)
final_f.sitetr1.unwunifrac <- f.sitetr1.unwunifrac/n_iterations

#SITE OF TRANSFER
final_pvalue.sitetr2.unwunifrac <- pvalue.sitetr2.unwunifrac/n_iterations
final_sumofsqs.sitetr2.1.unwunifrac <- sumofsqs.sitetr2.1.unwunifrac/n_iterations
final_sumofsqs.sitetr2.res.unwunifrac <- sumofsqs.sitetr2.res.unwunifrac/n_iterations
calculated_ss.sitetr2.total.unwunifrac <- final_sumofsqs.sitetr2.1.unwunifrac+final_sumofsqs.sitetr2.res.unwunifrac
final_r2.sitetr2.1.unwunifrac <- r2.sitetr2.unwunifrac/n_iterations
calculated_r2.sitetr2.res.unwunifrac = (1-final_r2.sitetr2.1.unwunifrac)
final_f.sitetr2.unwunifrac<- f.sitetr2.unwunifrac/n_iterations

#GROUP DISPERSION
final_pvalue.betadisper.sitetr1.unwunifrac <- pvalue.betadisper.sitetr1.unwunifrac/n_iterations #0.2114821
final_pvalue.betadisper.sitetr2.unwunifrac <- pvalue.betadisper.sitetr2.unwunifrac/n_iterations #0.827527

#MAKE TABLES
df_som_unwunifrac <- data.frame(var=c('sitetr1', 'sitetr2', 'Residual', 'Total'),
                                Df=c(1,1,30,32),
                                SumOfSqs=c(final_ss.som.sitetr1.unwunifrac,final_ss.som.sitetr2.unwunifrac,final_ss.som.res.unwunifrac,calculated_ss.som.total.unwunifrac),
                                R2=c(final_r2.som.sitetr1.unwunifrac,final_r2.som.sitetr2.unwunifrac,calculated_r2.som.res.unwunifrac,1),
                                "F"=c(final_f.1, final_f.2.unwunifrac ,"",""),
                                "p"=c(final_pvalue.adonis2.som.sitetr1.unwunifrac,final_pvalue.adonis2.som.sitetr2.unwunifrac,"",""))

df_sitetr1_unwunifrac <- data.frame(var=c('sitetr1', 'Residual', 'Total'),
                                    Df=c(1,31,32),
                                    SumOfSqs=c(final_sumofsqs.sitetr1.1.unwunifrac , final_sumofsqs.sitetr1.res.unwunifrac , calculated_ss.sitetr1.total.unwunifrac),
                                    R2=c(final_r2.som.sitetr1.unwunifrac , calculated_r2.sitetr1.res.unwunifrac , "1"),
                                    "F"=c(final_f.sitetr1.unwunifrac,"",""),
                                    "p"=c(final_pvalue.sitetr1.unwunifrac,"",""))

df_sitetr2_unwunifrac <- data.frame(var=c('sitetr2', 'Residual', 'Total'),
                                    Df=c(1,31,32),
                                    SumOfSqs=c(final_sumofsqs.sitetr2.1.unwunifrac , final_sumofsqs.sitetr2.res.unwunifrac , calculated_ss.sitetr2.total.unwunifrac),
                                    R2=c(final_r2.som.sitetr2.unwunifrac , calculated_r2.sitetr2.res.unwunifrac , "1"),
                                    "F"=c(final_f.sitetr2.unwunifrac,"",""),
                                    "p"=c(final_pvalue.sitetr2.unwunifrac,"",""))

#################################################
```
**Output table treatment**
```{r}
df_som_unwunifrac
write.table(df_som_unwunifrac, "adonis2_som_unwunifrac.txt")
```
**Output table in terms of site of origin**
```{r}
df_sitetr1_unwunifrac
write.table(df_sitetr1_unwunifrac, "adonis2_sitetr1_unwunifrac.txt")
```
**Betadisper p-value site of origin**
```{r}
final_pvalue.betadisper.sitetr1.unwunifrac
```
**Output table in terms of site of transfer**
```{r}
df_sitetr2_unwunifrac
write.table(df_sitetr2_unwunifrac, "adonis2_sitetr2_unwunifrac.txt")
```
**Betadisper p-value site of transfer**
```{r}
final_pvalue.betadisper.sitetr2.unwunifrac
```

### All metrics in terms of genetic cluster

```{r}
#ALPHA DIVERSITY
set.seed(123)
model3.geneticcluster<-lm(observed_features_log~geneticcluster, data=metadata.pup)
summary(model3.geneticcluster) #Adjusted R2 = 3256, p-value = 0.01551
set.seed(123)
model3.geneticcluster.sh<-lm(shannon_entropy~geneticcluster, data=metadata.pup)
summary(model3.geneticcluster.sh) #Adjusted R2 = 0.3214, p-value = 0.01667
set.seed(123)
model3.geneticcluster.pd<-lm(faith_pd~geneticcluster, data=metadata.pup)
summary(model3.geneticcluster.pd) #Adjusted R2 = 0.4175, p-value = 0.002517

#BETA DIVERSITY
set.seed(123)
bray_dist.pup = phyloseq::distance(psrt.pup, method="bray")
set.seed(123)
jacc_dist.pup = phyloseq::distance(psrt.pup, method="jaccard")
set.seed(123)
uu_dist.pup = phyloseq::distance(psrt.pup, method="uunifrac")
set.seed(123)
wu_dist.pup = phyloseq::distance(psrt.pup, method="wunifrac")
metadata.psrt.pup.gc <- as(sample_data(psrt.pup), "data.frame")
metadata.psrt.pup.gc$geneticcluster<-as.factor(metadata.psrt.pup.gc$geneticcluster)

set.seed(123)
pmod.pup.geneticcluster.bray<-adonis2(bray_dist.pup~geneticcluster, data=metadata.psrt.pup.gc, by="margin", perm=9999)
pmod.pup.geneticcluster.bray
set.seed(123)
pmod.pup.geneticcluster.jacc<-adonis2(jacc_dist.pup~geneticcluster, data=metadata.psrt.pup.gc, by="margin", perm=9999)
pmod.pup.geneticcluster.jacc
set.seed(123)
pmod.pup.geneticcluster.wu<-adonis2(wu_dist.pup~geneticcluster, data=metadata.psrt.pup.gc, by="margin", perm=9999)
pmod.pup.geneticcluster.wu
set.seed(123)
pmod.pup.geneticcluster.uu<-adonis2(uu_dist.pup~geneticcluster, data=metadata.psrt.pup.gc, by="margin", perm=9999)
pmod.pup.geneticcluster.uu
```



## Recreate Figure S1

**Individual taxa bar plots displaying the bacterial families in the post-transfer gut microbiota in adult and juvenile bank voles arranged according to experimental group**

```{r}
physeq2.f.3 = filter_taxa(psrt.mom.time3, function(x) mean(x) > 0.1, TRUE)
physeq3.f.3 = transform_sample_counts(physeq2.f.3, function(x) x / sum(x) )
glom.f.3 <- tax_glom(physeq3.f.3, taxrank = 'Family')
data_glom.3 <- psmelt(glom.f.3) 
data_glom.3$Family <- as.character(data_glom.3$Family) 
data_glom.3$Family <- factor(data_glom.3$Family, 
                             levels = c("NA","Campylobacteraceae",
                                        "Ruminococcaceae", "Christensenellaceae","Desulfovibrionaceae",
                                        "Oscillospiraceae","Lachnospiraceae",
                                        "Lactobacillaceae","Muribaculaceae"))

data_glom.3$treatment[data_glom.3$treatment=="urburb"]<-"Urban-Urban"
data_glom.3$treatment[data_glom.3$treatment=="urbcont"]<-"Urban-Rural"
data_glom.3$treatment[data_glom.3$treatment=="conturb"]<-"Rural-Urban"
data_glom.3$treatment[data_glom.3$treatment=="contcont"]<-"Rural-Rural"
plot.fig.S3<-ggplot(data_glom.3, 
       aes(x=Sample, y=Abundance, fill=Family)) + 
  geom_bar(stat="identity", position="fill",color="black") + 
  facet_grid(~treatment, scale="free")+
  scale_fill_manual(values = c("#9E0142", "#D53E4F", "#F46D43", "#FDAE61", "#FEE08B","#ABDDA4","blue","purple")) +
  theme_bw()+
  xlab("Individual adult bank vole")+
  ylab("Relative Abundance")+
  theme(axis.text.x=element_blank())

physeq2.f.pup = filter_taxa(psrt.pup, function(x) mean(x) > 0.1, TRUE)
physeq3.f.pup = transform_sample_counts(physeq2.f.pup, function(x) x / sum(x) )
glom.f.pup <- tax_glom(physeq3.f.pup, taxrank = 'Family')
data_glom.pup <- psmelt(glom.f.pup) 
data_glom.pup$Family <- as.character(data_glom.pup$Family) 
data_glom.pup$Family <- factor(data_glom.pup$Family, 
                             levels = c("NA","Campylobacteraceae",
                                        "Ruminococcaceae", "Christensenellaceae","Desulfovibrionaceae",
                                        "Oscillospiraceae","Lachnospiraceae",
                                        "Lactobacillaceae","Muribaculaceae"))

data_glom.pup$treatment[data_glom.pup$treatment=="urburb"]<-"Urban-Urban"
data_glom.pup$treatment[data_glom.pup$treatment=="urbcont"]<-"Urban-Rural"
data_glom.pup$treatment[data_glom.pup$treatment=="conturb"]<-"Rural-Urban"
data_glom.pup$treatment[data_glom.pup$treatment=="contcont"]<-"Rural-Rural"
plot.fig.S3.pup<-ggplot(data_glom.pup, 
                    aes(x=Sample, y=Abundance, fill=Family)) + 
  geom_bar(stat="identity", position="fill",color="black") + 
  facet_grid(~treatment, scale="free")+
  scale_fill_manual(values = c("#9E0142", "#D53E4F", "#F46D43", "#FDAE61", "#FEE08B","#ABDDA4","blue","purple")) +
  theme_bw()+
  xlab("Individual pup")+
  ylab("Relative Abundance")+
  theme(axis.text.x=element_blank())

plot.fig.S3.mom.pup<-ggarrange(plot.fig.S3, plot.fig.S3.pup,
                       common.legend = TRUE,
                       legend = "right",
                       labels = c("A", "B"),
                       ncol = 1, nrow = 2)

plot.fig.S3.mom.pup # 8 x 6

```

## Recreate Figure S2

**PCoA ordinations visualising the compositions of the pre-transfer gut microbiota of adult bank voles**

```{r}
#| fig-width: 9
#| fig-height: 6

p2.ord.bray.mom.1 <- ordinate(psrt.mom.time1, "PCoA", distance=bray_dist.mom.time1) 
bray.mom.1 = plot_ordination(psrt.mom.time1, p2.ord.bray.mom.1, type="samples", color="sitetr1full",  axes=c(1,2)) +
  stat_ellipse(aes(color=sitetr1), alpha=.5,linewidth =2)+
  scale_color_manual(values=c("#693aa3","green", "#f00078"), name=c(""),
                     labels=c("Rural (Jyvskyl)","Rural (Konnevesi)", "Urban"))+
  xlab("PCoA Axis.1  [16%]")+
  ylab("PCoA Axis.2  [9.6%]")+
  ggtitle("Bray-Curtis: Pre-transfer") +
  scale_y_continuous(position = "left")+
  theme_minimal() 

p2.ord.jacc.mom.1 <- ordinate(psrt.mom.time1, "PCoA", distance=jacc_dist.mom.time1) 
jacc.mom.1 = plot_ordination(psrt.mom.time1, p2.ord.jacc.mom.1, type="samples", color="sitetr1full",  axes=c(1,2)) +
  stat_ellipse(aes(color=sitetr1), alpha=.5,linewidth =2)+
  scale_color_manual(values=c("#693aa3","green", "#f00078"), name=c(""),
                     labels=c("Rural (Jyvskyl)","Rural (Konnevesi)", "Urban"))+
  xlab("PCoA Axis.1  [11%]")+
  ylab("PCoA Axis.2  [7.2%]")+
  ggtitle("Jaccard: Pre-transfer") +
  scale_y_continuous(position = "left")+
  theme_minimal() 

p2.ord.wunifrac.mom.1 <- ordinate(psrt.mom.time1, "PCoA", distance=wunifrac_dist.mom.time1) 
wunifrac.mom.1 = plot_ordination(psrt.mom.time1, p2.ord.wunifrac.mom.1, type="samples", color="sitetr1full",  axes=c(1,2)) +
  stat_ellipse(aes(color=sitetr1), alpha=.5,linewidth =2)+
  scale_color_manual(values=c("#693aa3","green", "#f00078"), name=c(""),
                     labels=c("Rural (Jyvskyl)","Rural (Konnevesi)", "Urban"))+
  xlab("PCoA Axis.1  [45.2%]")+
  ylab("PCoA Axis.2  [12%]")+
  ggtitle("wUnifrac: Pre-transfer") +
  scale_y_continuous(position = "left")+
  theme_minimal() 

p2.ord.uunifrac.mom.1 <- ordinate(psrt.mom.time1, "PCoA", distance=uunifrac_dist.mom.time1) 
uunifrac.mom.1 = plot_ordination(psrt.mom.time1, p2.ord.uunifrac.mom.1, type="samples", color="sitetr1full",  axes=c(1,2)) +
  stat_ellipse(aes(color=sitetr1), alpha=.5,linewidth =2)+
  scale_color_manual(values=c("#693aa3","green", "#f00078"), name=c(""),
                     labels=c("Rural (Jyvskyl)","Rural (Konnevesi)", "Urban"))+
  xlab("PCoA Axis.1  [10.2%]")+
  ylab("PCoA Axis.2  [9.8%]")+
  ggtitle("unwUnifrac: Pre-transfer") +
  scale_y_continuous(position = "left")+
  theme_minimal() 

plot.fig.t1<-ggarrange(bray.mom.1, jacc.mom.1,wunifrac.mom.1,uunifrac.mom.1,
                       common.legend = TRUE,
                       legend = "right",
                       labels = c("A", "B","C","D"),
                       ncol = 2, nrow = 2)

plot.fig.t1 # 7 x 6
```

## Recreate Figure S3

**Change in alpha diversity of the gut microbiota of bank voles throughout the experiment (ASV Richness)**

```{r}
#| fig-width: 7
#| fig-height: 4
plot.asv.t1.to.t3<-ggplot(data=diff.13, aes(x=treatment, y=asv_1_to_3, fill=treatment))+
  geom_boxplot() +
  theme(legend.position="right")+
  labs(title= "ASV richness", x= "",y = "Change in alpha diversity")+
  scale_x_discrete(labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  scale_fill_manual(values=c("#693aa3","#7777ff","#ffc1e0","#f00078"),name=c("RT exp. group"),
                    labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  theme_minimal()

plot.asv.t3<-ggplot(data=metadata.mom.time3, aes(x=treatment, y=observed_features, fill=treatment))+
  geom_boxplot() +
  theme(legend.position="right")+
  labs(title= "Post-transfer ASV richness", x= "",y = "Total alpha diversity")+
  scale_x_discrete(labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  scale_fill_manual(values=c("#693aa3","#7777ff","#ffc1e0","#f00078"),name=c("RT exp. group"),
                    labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  theme_minimal()

plot.alpha.change<-ggarrange(plot.asv.t1.to.t3+ rremove("x.text"), plot.asv.t3 + rremove("x.text"),
                             common.legend = TRUE,
                             legend = "right",
                             labels = c("A", "B"),
                             ncol = 2, nrow = 1)

plot.alpha.change 
```

## Recreate Figure S4

**Magnitude of change in beta diversity of the gut microbiota of bank voles in terms of past habitat**

```{r}
plot.bray.t1.to.t3<-ggplot(data=dist.13, aes(x=sitetr1, y=bray_1_to_3, fill=sitetr1))+
  geom_boxplot() +
  theme(legend.position="right")+
  labs(title= "Bray-Curtis metric", x= "",y = "Change in Beta Diversity")+
  scale_x_discrete(labels=c("Rural","Urban"))+
  scale_fill_manual(values=c("#693aa3","#f00078"),name=c(""),
                    labels=c("Rural","Urban"))+
  geom_signif(y_position = c(0.95), xmin = c(1), xmax = c(2),annotation = c("ns"), tip_length = 0)+
  theme_minimal()

plot.jacc.t1.to.t3<-ggplot(data=dist.13, aes(x=sitetr1, y=jacc_1_to_3, fill=sitetr1))+
  geom_boxplot() +
  theme(legend.position="right")+
  labs(title= "Jaccard metric", x= "",y = "Change in Beta Diversity")+
  scale_x_discrete(labels=c("Rural","Urban"))+
  scale_fill_manual(values=c("#693aa3","#f00078"),name=c(""),
                    labels=c("Rural","Urban"))+
  geom_signif(y_position = c(0.875), xmin = c(1), xmax = c(2),annotation = c("ns"), tip_length = 0)+
  theme_minimal()

plot.wunifrac.t1.to.t3<-ggplot(data=dist.13, aes(x=sitetr1, y=wunifrac_1_to_3, fill=sitetr1))+
  geom_boxplot() +
  theme(legend.position="right")+
  labs(title= "wUnifrac metric", x= "",y = "Change in Beta Diversity")+
  scale_x_discrete(labels=c("Rural","Urban"))+
  scale_fill_manual(values=c("#693aa3","#f00078"),name=c(""),
                    labels=c("Rural","Urban"))+
  geom_signif(y_position = c(0.45), xmin = c(1), xmax = c(2),annotation = c("*"), tip_length = 0)+
  theme_minimal()

plot.unwunifrac.t1.to.t3<-ggplot(data=dist.13, aes(x=sitetr1, y=unwunifrac_1_to_3, fill=sitetr1))+
  geom_boxplot() +
  theme(legend.position="right")+
  labs(title= "unwUnifrac metric", x= "",y = "Change in Beta Diversity")+
  scale_x_discrete(labels=c("Rural","Urban"))+
  scale_fill_manual(values=c("#693aa3","#f00078"),name=c(""),
                    labels=c("Rural","Urban"))+
  geom_signif(y_position = c(0.575), xmin = c(1), xmax = c(2),annotation = c("ns"), tip_length = 0)+
  theme_minimal()

plot.fig.turnover<-ggarrange(plot.bray.t1.to.t3+ rremove("x.text"), plot.jacc.t1.to.t3 + rremove("x.text"),plot.wunifrac.t1.to.t3+ rremove("x.text"),plot.unwunifrac.t1.to.t3+ rremove("x.text"),
                             common.legend = TRUE,
                             legend = "right",
                             labels = c("A", "B","C","D"),
                             ncol = 2, nrow = 2)

plot.fig.turnover 
```

## Recreate Figure S5

**Additional CAP ordinations visualising the compositions of adult and juvenile post-transfer gut microbiota in terms of past and present habitat**

```{r}
#| fig-width: 12
#| fig-height: 10

#######
#ADULTS
#######
set.seed(123)
cap.mom.3.jacc <- ordinate(physeq = psrt.mom.time3, method = "CAP",distance = jacc_dist.mom.time3,formula = ~ sitetr1+sitetr2)
plot.mom.3.jacc<-plot_ordination(psrt.mom.time3, cap.mom.3.jacc,type = "samples",color="treatment")+geom_point(size = 3)+
  scale_color_manual(values=c("#693aa3","#7777ff","#ffc1e0","#f00078"), name=(""),
                     labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  ggtitle("Adults: Jaccard metric")+xlab("CAP Axis.1  [6.4%]")+ylab("CAP Axis.2  [3.7%]")+theme_minimal()

set.seed(123)
cap.mom.3.wunifrac <- ordinate(physeq = psrt.mom.time3, method = "CAP",distance = wunifrac_dist.mom.time3,formula = ~ sitetr1+sitetr2)
plot.mom.3.wunifrac<-plot_ordination(psrt.mom.time3, cap.mom.3.wunifrac,type = "samples",color="treatment")+geom_point(size = 3)+
  scale_color_manual(values=c("#693aa3","#7777ff","#ffc1e0","#f00078"), name=(""),
                     labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  ggtitle("Adults: wUnifrac metric")+xlab("CAP Axis.1  [7.9%]")+ylab("CAP Axis.2  [4.5%]")+theme_minimal()

set.seed(123)
cap.mom.3.unwunifrac <- ordinate(physeq = psrt.mom.time3, method = "CAP",distance = uunifrac_dist.mom.time3,formula = ~ sitetr1+sitetr2)
plot.mom.3.uunifrac<-plot_ordination(psrt.mom.time3, cap.mom.3.unwunifrac,type = "samples",color="treatment")+geom_point(size = 3)+
  scale_color_manual(values=c("#693aa3","#7777ff","#ffc1e0","#f00078"), name=(""),
                     labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  ggtitle("Adults: unwUnifrac metric")+xlab("CAP Axis.1  [4.8%]")+ylab("CAP Axis.2  [4.4%]")+theme_minimal()

plot.cap.adults.si<-ggarrange(plot.mom.3.jacc, plot.mom.3.wunifrac,plot.mom.3.uunifrac,
                              common.legend = TRUE,
                              legend = "right",
                              labels = c("A","B", "C"),
                              ncol = 3, nrow = 1, align="v"+ theme_void())

########
#OFFSPRING
########
set.seed(123)
jacc_dist.pup.all = phyloseq::distance(psrt.pup, method="jaccard")
set.seed(123)
wunifrac_dist.pup.all = phyloseq::distance(psrt.pup, method="wunifrac")
set.seed(123)
uunifrac_dist.pup.all = phyloseq::distance(psrt.pup, method="uunifrac")
metadata.psrt.pup.all <- as(sample_data(psrt.pup), "data.frame")
set.seed(123)
cap.pup.all.jacc <- ordinate(physeq = psrt.pup, method = "CAP",distance = jacc_dist.pup.all,formula = ~ sitetr1+sitetr2)
set.seed(123)
cap.pup.all.wunifrac <- ordinate(physeq = psrt.pup, method = "CAP",distance = wunifrac_dist.pup.all,formula = ~ sitetr1+sitetr2)
set.seed(123)
cap.pup.all.unwunifrac <- ordinate(physeq = psrt.pup, method = "CAP",distance = uunifrac_dist.pup.all,formula = ~ sitetr1+sitetr2)

plot.pup.cap.jacc<-plot_ordination(psrt.pup, cap.pup.all.jacc,type = "samples",color="treatment")+geom_point(size = 3)+
  scale_color_manual(values=c("#693aa3","#7777ff","#ffc1e0","#f00078"), name=(""),
                     labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  ggtitle("Offspring: Jaccard metric")+xlab("CAP Axis.1  [3%]")+ylab("CAP Axis.2  [1.8%]")+theme_minimal()

plot.pup.cap.wunifrac<-plot_ordination(psrt.pup, cap.pup.all.wunifrac,type = "samples",color="treatment")+geom_point(size = 3)+
  scale_color_manual(values=c("#693aa3","#7777ff","#ffc1e0","#f00078"), name=(""),
                     labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  ggtitle("Offspring: wUnifrac metric")+xlab("CAP Axis.1  [2.9%]")+ylab("CAP Axis.2  [1.9%]")+theme_minimal()

plot.pup.cap.unwunifrac<-plot_ordination(psrt.pup, cap.pup.all.unwunifrac,type = "samples",color="treatment")+geom_point(size = 3)+
  scale_color_manual(values=c("#693aa3","#7777ff","#ffc1e0","#f00078"), name=(""),
                     labels=c("Rural-Rural","Rural-Urban","Urban-Rural","Urban-Urban"))+
  ggtitle("Offspring: unwUnifrac metric")+xlab("CAP Axis.1  [3%]")+ylab("CAP Axis.2  [2.5%]")+theme_minimal()

plot.cap.offspring.si<-ggarrange(plot.pup.cap.jacc, plot.pup.cap.wunifrac,plot.pup.cap.unwunifrac,
                              common.legend = TRUE,
                              legend = "right",
                              labels = c("D","E", "F"),
                              ncol = 3, nrow = 1, align="v"+ theme_void())

plot.cap.all.si<-ggarrange(plot.cap.adults.si, plot.cap.offspring.si,
                              common.legend = TRUE,
                              legend = "right",
                              ncol = 1, nrow = 2, align="v"+ theme_void())
plot.cap.all.si # 10 x 12

```

## Recreate Figure S6

**PCoA ordinations visualising the compositions of adult and juvenile post-transfer gut microbiota in terms of genetic cluster**

```{r}
#| fig-width: 10
#| fig-height: 5
set.seed(123)
bray_dist.pup = phyloseq::distance(psrt.pup, method="bray")
set.seed(123)
jacc_dist.pup = phyloseq::distance(psrt.pup, method="jaccard")
set.seed(123)
uu_dist.pup = phyloseq::distance(psrt.pup, method="uunifrac")
set.seed(123)
wu_dist.pup = phyloseq::distance(psrt.pup, method="wunifrac")

#MATCHED mothers + offspring
set.seed(123)
bray_dist.matched = phyloseq::distance(psrt.matched, method="bray")
set.seed(123)
jacc_dist.matched = phyloseq::distance(psrt.matched, method="jaccard")
set.seed(123)
uunifrac_dist.matched = phyloseq::distance(psrt.matched, method="uunifrac")
set.seed(123)
wunifrac_dist.matched = phyloseq::distance(psrt.matched, method="wunifrac")

p2.ord.bray.matched <- ordinate(psrt.matched, "PCoA", distance=bray_dist.matched) 
bray.matched = plot_ordination(psrt.matched, p2.ord.bray.matched, type="samples", color="geneticcluster_2",  shape="type",axes=c(1,2)) +
   geom_point(alpha=2, size=3)+
  labs(color='Genetic cluster',shape="Type")+ 
  xlab("PCoA Axis.1  [15.5%]")+
  ylab("PCoA Axis.2  [10.7%]")+
  ggtitle("Matched mothers and offspring (Bray-Curtis metric)") +
  scale_y_continuous(position = "left")+
  theme_minimal() 

p2.ord.bray.pup <- ordinate(psrt.pup, "PCoA", distance=bray_dist.pup) 
plot.pcoa.pup.bray = plot_ordination(psrt.pup, p2.ord.bray.pup, type="samples", color="geneticcluster_2",axes=c(1,2)) +
  geom_point(alpha=2, size=3)+
  labs(color='Genetic cluster')+ 
  xlab("PCoA Axis.1  [11.3%]")+
  ylab("PCoA Axis.2  [9.5%]")+
  ggtitle("All offspring (Bray-Curtis metric)") +
  scale_y_continuous(position = "left")+
  theme_minimal() 

plot.geneticcluster<-ggarrange(bray.matched, plot.pcoa.pup.bray,
                              common.legend = FALSE,
                              legend = "right",
                              labels = c("A","B"),
                              ncol = 2, nrow = 1, align="v"+ theme_void())
plot.geneticcluster # 10 x 12


```
